---
title: "AIAttitudes Code"
author: "Matthew Jones"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    number_sections: true
---

```{r setup, include=FALSE}
library(tidyverse) 
pacman::p_load(haven, survey, car, ggplot2, ggthemes, skimr, ggthemes, ggpubr, gridExtra, tidyr, dplyr, gt, sjlabelled, RColorBrewer, polycor, corrplot, paletteer, psych, kableExtra, patchwork, GPArotation, stargazer, poLCA, reshape2)

knitr::opts_chunk$set(echo = TRUE)
options(tinytex.verbose = TRUE)
```

# Dataset

International Social Survey Programme (ISSP) & Purhonen, Semi (Tampereen yliopisto): ISSP 2024: digitaalinen yhteiskunta: Suomen aineisto [data]. Dataversio 3.0 (2026-01-18). Yhteiskuntatieteellinen tietoarkisto [jakaja]. DOI: https://doi.org/10.60686/t-fsd3994; URN: https://urn.fi/urn:nbn:fi:fsd:T-FSD3994


# Data wrangling

1.  All value scales in this report are organized from low to high. When necessary, variables from the original data set have been reverse coded and have been labeled with an "x" (for example, the rescaled k32a is identified as k32ax).

2.  All questions and value scales have been translated to English from the original Finnish language with assistance from "Google Translate". Where the software translation was unclear or imprecise, consultations with native Finnish speakers were employed to ensure clarity and meaning.

```{r Data Import, eval=FALSE, include=FALSE}
# Import data to dataframe
daF3994_fin <- read_sav("./Data/daF3994_fin.sav")

```


```{r Temp}
daF3994_fin <- read_sav("Datasets/FSD3994/Study/Data/daF3994_fin.sav")
```

```{r Data wrangling1}
# Select study variables
AiR_base <- subset(daF3994_fin, select= c(fsd_id, k1, ika, k3, k4, k70, nuts2, k23, k32a, k32b, k32c, k32d, k34a, k34b, k46, k47, k48a, k48b, k57, k62, k58a, k58b, k58c, k58d, k58e, otospaino, analyysipaino))

skim(AiR_base)
```

```{r Data wrangling2 reorder low to high, include=FALSE}
# Reorder categorical variables from low to high
AiR_base$k23x <- 8 - AiR_base$k23
AiR_base$k32ax <- 6 - AiR_base$k32a
AiR_base$k32bx <- 6 - AiR_base$k32b
AiR_base$k32cx <- 6 - AiR_base$k32c
AiR_base$k32dx <- 6 - AiR_base$k32d
AiR_base$k34ax <- 6 - AiR_base$k34a
AiR_base$k34bx <- 6 - AiR_base$k34b
AiR_base$k57x <- 6 - AiR_base$k57

```

```{r Data wrangling3 label variables}
# Add labels to study variables
AiR_base$fsd_id <- set_label(AiR_base$fsd_id, label = "Observation ID")
AiR_base$k1 <- set_label(AiR_base$k1, label = "Gender")
AiR_base$ika <- set_label(AiR_base$ika, label = "Age")
AiR_base$k3 <- set_label(AiR_base$k3, label = "Education (years)")
AiR_base$k4 <- set_label(AiR_base$k4, label = "Education level")
AiR_base$k70 <- set_label(AiR_base$k70, label = "Marital Status")
AiR_base$nuts2 <- set_label(AiR_base$nuts2, label = "Region")
AiR_base$k23x <- set_label(AiR_base$k23x, label = "Internet use (level)")
AiR_base$k32ax <- set_label(AiR_base$k32ax, label = "Ability level: Using internet")
AiR_base$k32bx <- set_label(AiR_base$k32bx, label = "Ability level: Searching online")
AiR_base$k32cx <- set_label(AiR_base$k32cx, label = "Ability level: Online privacy")
AiR_base$k32dx <- set_label(AiR_base$k32dx, label = "Ability level: Learning new software")
AiR_base$k34ax <- set_label(AiR_base$k34ax, label = "It is more comfortable meeting online")
AiR_base$k34bx <- set_label(AiR_base$k34bx, label = "Without internet, I would feel lonely")
AiR_base$k46 <- set_label(AiR_base$k46, label = "Political leaning")
AiR_base$k47 <- set_label(AiR_base$k47, label = "Trust level: People in general")
AiR_base$k48a <- set_label(AiR_base$k48a, label = "Trust level: Finnish parliament")
AiR_base$k48b <- set_label(AiR_base$k48b, label = "Trust level: Finnish courts")
AiR_base$k62 <- set_label(AiR_base$k62, label = "Social level: Self-assessed")
AiR_base$k57x <- set_label(AiR_base$k57x, label = "Worry that AI/Computer/Machine will take jobs")
AiR_base$k58a <- set_label(AiR_base$k58a, label = "Comfort level: AI in medical procedure")
AiR_base$k58b <- set_label(AiR_base$k58b, label = "Comfort level: AI self-driving cars")
AiR_base$k58c <- set_label(AiR_base$k58c, label = "Comfort level: Human/AI uncertainty")
AiR_base$k58d <- set_label(AiR_base$k58d, label = "Comfort level: Communication with AI")
AiR_base$k58e <- set_label(AiR_base$k58e, label = "Comfort level: Reliability of AI information")
AiR_base$otospaino <- set_label(AiR_base$otospaino, label = "Sample weight")
AiR_base$analyysipaino <- set_label(AiR_base$analyysipaino, label = "Analysis weight")

```

```{r Data wrangling4 build study set}
# Build a study set that includes dbl and factor columns
AiR_num <- subset(AiR_base, select= c(fsd_id, k1, ika, k3, k4, k70, nuts2, k23x, k32ax, k32bx, k32cx, k32dx, k34ax, k34bx, k46, k47, k48a, k48b, k57x, k62, k58a, k58b, k58c, k58d, k58e, otospaino, analyysipaino))

# Specify the variables to convert
vars_to_convert <- c("k1", "k4", "k70", "nuts2", "k23x", "k32ax", 
                     "k32bx", "k32cx", "k32dx", "k34ax", "k34bx", 
                     "k46", "k47", "k48a", "k48b", "k57x", "k62", 
                     "k58a", "k58b", "k58c", "k58d", "k58e")

# Mutate the specified variables into factors
AiR_cat <- AiR_num %>%
  mutate(across(all_of(vars_to_convert), as.factor))


# Apply labels to factor variables
AiR_cat$k1 <- set_label(AiR_cat$k1, label = "Gender")
AiR_cat$k4 <- set_label(AiR_cat$k4, label = "Education level")
AiR_cat$k70 <- set_label(AiR_cat$k70, label = "Marital Status")
AiR_cat$nuts2 <- set_label(AiR_cat$nuts2, label = "Region")
AiR_cat$k23x <- set_label(AiR_cat$k23x, label = "Internet use (level)")
AiR_cat$k32ax <- set_label(AiR_cat$k32ax, label = "Ability level: Using internet")
AiR_cat$k32bx <- set_label(AiR_cat$k32bx, label = "Ability level: Searching online")
AiR_cat$k32cx <- set_label(AiR_cat$k32cx, label = "Ability level: Online privacy")
AiR_cat$k32dx <- set_label(AiR_cat$k32dx, label = "Ability level: Learning new software")
AiR_cat$k34ax <- set_label(AiR_cat$k34ax, label = "It is more comfortable meeting online")
AiR_cat$k34bx <- set_label(AiR_cat$k34bx, label = "Without internet, I would feel lonely")
AiR_cat$k46 <- set_label(AiR_cat$k46, label = "Political leaning")
AiR_cat$k47 <- set_label(AiR_cat$k47, label = "Trust level: People in general")
AiR_cat$k48a <- set_label(AiR_cat$k48a, label = "Trust level: Finnish parliament")
AiR_cat$k48b <- set_label(AiR_cat$k48b, label = "Trust level: Finnish courts")
AiR_cat$k62 <- set_label(AiR_cat$k62, label = "Social level: Self-assessed")
AiR_cat$k57x <- set_label(AiR_cat$k57x, label = "Worry that AI will take jobs")
AiR_cat$k58a <- set_label(AiR_cat$k58a, label = "Comfort level: AI in medical procedure")
AiR_cat$k58b <- set_label(AiR_cat$k58b, label = "Comfort level: AI self-driving cars")
AiR_cat$k58c <- set_label(AiR_cat$k58c, label = "Comfort level: Human/AI uncertainty")
AiR_cat$k58d <- set_label(AiR_cat$k58d, label = "Comfort level: Communication with AI")
AiR_cat$k58e <- set_label(AiR_cat$k58e, label = "Comfort level: Reliability of AI information")

# Append a factor identifier to var names
AiR_fac <- AiR_cat %>%
  rename_with(~ paste0(.x, "_f"), everything())

# Build new dataframe with numeric and factor copies of vars
AiR <- cbind(AiR_num, AiR_fac)

```

```{r Data wrangling5 value label vectors}
# Vector for gender
gen3lik <- c(
  `1` = "Male",
  `2` = "Female",
  `3` = "Prefer not to say"
  ) 

# Vector for education
edu13lik <- c(
  `0` = "Haven't gone to school",
  `1` =	"Still in comprehensive school",
  `2` = "Elementary school or civic school",
  `3` = "Comprehensive school",
  `4` = "Trade school or course",
  `5` = "High school or matriculation examination",
  `6` = "Specialist Vocational Qualification",
  `7` = "Post-secondary vocational education",
  `8` = "Bachelor's degree",
  `9` = "Master's Degree",
  `10` = "University, bachelor's degree",
  `11` = "University, master's degree",
  `12` = "Licentiate or doctoral degree"
  ) 


# Vector for marital status
mar6lik <- c(
  `1` =	"Married",
  `2` =	"In a registered partnership",
  `3` =	"Separated (in a registered partnership)",
  `4` =	"Divorced from registered partnership",
  `5` =	"Widow after registered partnership",
  `6` =	"Single"
  )

# vector for region
reg5lik <- c(
  `1` =	"FI19 Western Finland",
  `2` =	"FI1B Helsinki-Uusimaa",
  `3` =	"FI1C Southern Finland",
  `4` =	"FI1D Northern and Eastern Finland",
  `5` =	"FI20 Åland Islands"
  )

freq7lik <- c(
  `1` = "Never",
  `2` = "Less often",
  `3` = "Several times a month",
  `4` = "Several times a week",
  `5` = "Once a day",
  `6` = "Several times a day",
  `7` = "All the time"
  ) 

# Vector for ability level
abil5lik <- c(
  `1` = "Very bad",
  `2` = "Bad",
  `3` = "Moderate",
  `4` = "Good",
  `5` = "Very good"
  ) 


agr5lik <- c(
  `1` = "Completely disagree",
  `2` = "Disagree",
  `3` = "Neither",
  `4` = "Agree",
  `5` = "Totally agree"
  ) 

#Vector for politics scale
pol11scal <- c(
  `0` = "0 (Left wing)",
  `1` =	"1",
  `2` = "2",
  `3` = "3",
  `4` = "4",
  `5` = "5 (Center)",
  `6` = "6",
  `7` = "7",
  `8` = "8",
  `9` = "9",
  `10` = "10 (Right wing)"
  ) 

#Vector for trust scale
trust11scal <- c(
  `0` = "0 (Completely distrust)",
  `1` =	"1",
  `2` = "2",
  `3` = "3",
  `4` = "4",
  `5` = "5",
  `6` = "6",
  `7` = "7",
  `8` = "8",
  `9` = "9",
  `10` = "10 (Trust completely)"
  ) 

# Vector for worry level
worry5lik <- c(
  `1` = "Not at all",
  `2` = "Not really",
  `3` = "A little",
  `4` = "Moderately",
  `5` = "Very"
  ) 

comf11scal <- c(
  `0` = "0 (Completely uncomfortable)",
  `1` =	"1",
  `2` = "2",
  `3` = "3",
  `4` = "4",
  `5` = "5",
  `6` = "6",
  `7` = "7",
  `8` = "8",
  `9` = "9",
  `10` = "10 (Completely comfortable)"
  )
```

```{r Data wrangling6 apply value labels, include=FALSE}

#Label factor values
AiR$k1_f <- set_labels(AiR$k1_f, labels = gen3lik)
AiR$k4_f <- set_labels(AiR$k4_f, labels = edu13lik)
AiR$k70_f <- set_labels(AiR$k70_f, labels = mar6lik)
AiR$nuts2_f <- set_labels(AiR$nuts2_f, labels = reg5lik)
AiR$k23x_f <- set_labels(AiR$k23x_f, labels = freq7lik)
AiR$k32ax_f <- set_labels(AiR$k32ax_f, labels = abil5lik)
AiR$k32bx_f <- set_labels(AiR$k32bx_f, labels = abil5lik)
AiR$k32cx_f <- set_labels(AiR$k32cx_f, labels = abil5lik)
AiR$k32dx_f <- set_labels(AiR$k32dx_f, labels = abil5lik)
AiR$k34ax_f <- set_labels(AiR$k34ax_f, labels = agr5lik)
AiR$k34bx_f <- set_labels(AiR$k34bx_f, labels = agr5lik)
AiR$k46_f <- set_labels(AiR$k46_f, labels = pol11scal)
AiR$k47_f <- set_labels(AiR$k47_f, labels = trust11scal)
AiR$k48a_f <- set_labels(AiR$k48a_f, labels = trust11scal)
AiR$k48b_f <- set_labels(AiR$k48b_f, labels = trust11scal)
AiR$k57x_f <- set_labels(AiR$k57x_f, labels = worry5lik)
AiR$k58a_f <- set_labels(AiR$k58a_f, labels = comf11scal)
AiR$k58b_f <- set_labels(AiR$k58b_f, labels = comf11scal)
AiR$k58c_f <- set_labels(AiR$k58c_f, labels = comf11scal)
AiR$k58d_f <- set_labels(AiR$k58d_f, labels = comf11scal)
AiR$k58e_f <- set_labels(AiR$k58e_f, labels = comf11scal)

```


# Outcome variables

```{r attitude plots, fig.cap="Distribution of attitudinal indictators", fig.height=6}
# k58a plot
k58a_plot <- ggplot(AiR, aes(x = k58a_f)) +
  geom_bar(fill="steelblue", color="black", alpha = 0.7) +
  labs(y = "Count") +
  theme_pubr()
# k58b plot
k58b_plot <- ggplot(AiR, aes(x = k58b_f)) +
  geom_bar(fill="steelblue", color="black", alpha = 0.7) +
  labs(y = "Count") +
  theme_pubr()
# k58c plot
k58c_plot <- ggplot(AiR, aes(x = k58c_f)) +
  geom_bar(fill="steelblue", color="black", alpha = 0.7) +
  labs(y = "Count") +
  theme_pubr()
# k58d plot
k58d_plot <- ggplot(AiR, aes(x = k58d_f)) +
  geom_bar(fill="steelblue", color="black", alpha = 0.7) +
  labs(y = "Count") +
  theme_pubr()
# k58e plot
k58e_plot <- ggplot(AiR, aes(x = k58e_f)) +
  geom_bar(fill="steelblue", color="black", alpha = 0.7) +
  labs(y = "Count") +
  theme_pubr()


attitude_plot <- wrap_plots(
  k58a_plot, k58b_plot, 
  k58c_plot, k58d_plot, 
  k58e_plot, 
  ncol = 2
) + 
  plot_annotation(caption = "Data from FSD3994 ISSP 2024: Digital Societies: Finnish Data")

attitude_plot + 
  theme(plot.margin = margin(10, 10, 10, 10))
```

## Test for evidence of straight-lining

```{r Staightlining sorting variable}

#Create indicator variable to report cases of all ones across AI indicators
AiR <- AiR %>% mutate(all1 = ifelse(k58a * k58b * k58c * k58d * k58e == 1, 1, 0)
                                  , all1 = as.factor(all1))
#Check variable count
summary(AiR$all1)

#Create subset for potential straight-liners
AiRstr <- AiR %>% subset(all1==1)
```


```{r Ind str test, fig.cap="Straight-line check across other indicators", fig.height=6}
# k32a plot
k32a_plot_s <- ggplot(AiRstr, aes(x = k32ax_f)) +
  geom_bar(fill="steelblue", color="black", alpha = 0.7) +
  labs(y = "Count",
       x = get_label(AiR$k32ax_f)) +
  theme_pubr()
# k32b plot
k32b_plot_s <- ggplot(AiRstr, aes(x = k32bx_f)) +
  geom_bar(fill="steelblue", color="black", alpha = 0.7) +
  labs(y = "Count",
       x = get_label(AiR$k32bx_f)) +
  theme_pubr()
# k32c plot
k32c_plot_s <- ggplot(AiRstr, aes(x = k32cx_f)) +
  geom_bar(fill="steelblue", color="black", alpha = 0.7) +
  labs(y = "Count",
       x = get_label(AiR$k32cx_f)) +
  theme_pubr()
# k32d plot
k32d_plot_s <- ggplot(AiRstr, aes(x = k32dx_f)) +
  geom_bar(fill="steelblue", color="black", alpha = 0.7) +
  labs(y = "Count",
       x = get_label(AiR$k32dx_f)) +
  theme_pubr()
k48a_plot_s <- ggplot(AiRstr, aes(x = k48a_f)) +
  geom_bar(fill="steelblue", color="black", alpha = 0.7) +
  labs(y = "Count",
       x = get_label(AiR$k48a_f)) +
  theme_pubr()
k57x_plot_s <- ggplot(AiRstr, aes(x = k57x_f)) +
  geom_bar(fill="steelblue", color="black", alpha = 0.7) +
  labs(y = "Count",
       x = get_label(AiR$k57x_f)) +
  theme_pubr()

digeff_plot <- wrap_plots(
  k32a_plot_s, k32b_plot_s, 
  k32c_plot_s, k32d_plot_s,
  k48a_plot_s, k57x_plot_s,
  ncol = 2
) + 
  plot_annotation(caption = "Data from FSD3994 ISSP 2024: Digital Societies: Finnish Data")

digeff_plot + 
  theme(plot.margin = margin(10, 10, 10, 10))
```

## Missingness analysis

```{r MCAR test variables 1}

AiR_ind <- subset(AiR, select= c(k1, ika, k3, k4, k32ax, k32bx, k32cx, k32dx, k48a, k57x, k58a, k58b, k58c, k58d, k58e))

na_counts <- sapply(AiR_ind[, c("k32ax", "k32bx", "k32cx", "k32dx", "k58a", "k58b", "k58c", "k58d", "k58e")], function(x) sum(is.na(x)))
print(na_counts)

AiR_test <- AiR %>% filter(if_all(c(k32ax, k32bx, k32cx, k32dx, k58a, k58b, k58c, k58d, k58e), ~ !is.na(.)))
missing_data <- AiR_ind %>%
  filter(is.na(k32ax) | is.na(k32bx) | is.na(k32cx) | is.na(k32dx) | is.na(k58a) | is.na(k58b) | is.na(k58c) | is.na(k58d) | is.na(k58e))
```

```{r MCAR test 1}
library(dplyr)

indicators <- c("k1", "ika", "k3", "k4", "k32ax", "k32bx", "k32cx", "k32dx", "k48a", "k57x", "k58a", "k58b", "k58c", "k58d", "k58e")

# Function to compare distribution in NA set to complete set
compare_distributions <- function(complete_data, subset_data, indicators) {
  results_df <- data.frame(Test_Variable = character(),
                            Statistic = numeric(),
                            P_value = numeric(),
                            stringsAsFactors = FALSE)  # Create an empty data frame for results

  for (indicator in indicators) {
    # Check if the indicator exists in both datasets
    if (indicator %in% colnames(complete_data) && indicator %in% colnames(subset_data)) {
      # Perform the Mann-Whitney U Test
      test_result <- wilcox.test(complete_data[[indicator]], subset_data[[indicator]], alternative = "two.sided")
      
      # Append the results to the data frame
      results_df <- rbind(results_df, data.frame(Indicator = indicator,
                                                  Statistic = test_result$statistic,
                                                  P_value = test_result$p.value,
                                                  stringsAsFactors = FALSE))
      
      # Print the results
      cat("Mann-Whitney U Test for", indicator, ":\n")
      cat("  Statistic:", test_result$statistic, "\n")
      cat("  P-value:", test_result$p.value, "\n\n")
    } else {
      cat("Indicator", indicator, "not found in one of the datasets.\n\n")
    }
  }
  
  return(results_df)  # Return the results data frame
}

# Call the function and store results in a variable
results_table <- compare_distributions(AiR_ind, AiR_test, indicators)

```

```{r MCAR table results, tab.cap="Mann-Whitney test for significance of removed observations"}

wilcox_table <- results_table %>%
  gt() %>%
  tab_header(
    title = "Results of Mann-Whitney U Test"
  ) %>%
  cols_label(
    Indicator = "Test Variable",
    Statistic = "Statistic",
    P_value = "P-value"
  ) %>%
  fmt_number(
    columns = c(Statistic, P_value),
    decimals = 2
  )
wilcox_table
```

Using the Mann-Whitney U test, the distributions of all indicator variables in the "clean" data set were compared to the original. For each variable, the test did not reveal significant cause to dismiss the null hypothesis that the distributions are from the same sample. Together, these results suggest that the missing data is likely missing completely at random (MCAR) and can be safely removed without skewing the results.

```{r Filter complete case analysis}
AiR_clean <- AiR %>%
  filter(if_all(c(k32ax, k32bx, k32cx, k32dx, k58a, k58b, k58c, k58d, k58e), ~ !is.na(.)))
```

\clearpage

# Predictor variables

## Digital technology efficacy

```{r digital eff plots, fig.cap="Distribution of digital efficacy indicators", fig.height=6}
# k32a plot
k32a_plot <- ggplot(AiR_clean, aes(x = k32ax_f)) +
  geom_bar(fill="steelblue", color="black", alpha = 0.7) +
  labs(y = "Count") +
  scale_x_discrete(labels = get_labels(AiR_clean$k32ax_f))+
  theme_pubr()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size=8))
# k32b plot
k32b_plot <- ggplot(AiR_clean, aes(x = k32bx_f)) +
  geom_bar(fill="steelblue", color="black", alpha = 0.7) +
  labs(y = "Count") +
  scale_x_discrete(labels = get_labels(AiR_clean$k32ax_f))+
  theme_pubr()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size=8))
# k32c plot
k32c_plot <- ggplot(AiR_clean, aes(x = k32cx_f)) +
  geom_bar(fill="steelblue", color="black", alpha = 0.7) +
  labs(y = "Count") +
  scale_x_discrete(labels = get_labels(AiR_clean$k32ax_f))+
  theme_pubr()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size=8))
# k32d plot
k32d_plot <- ggplot(AiR_clean, aes(x = k32dx_f)) +
  geom_bar(fill="steelblue", color="black", alpha = 0.7) +
  labs(y = "Count") +
  scale_x_discrete(labels = get_labels(AiR_clean$k32ax_f))+
  theme_pubr()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size=8))


digeff_plot <- wrap_plots(
  k32a_plot, k32b_plot, 
  k32c_plot, k32d_plot, 
  ncol = 2
) + 
  plot_annotation(caption = "Data from FSD3994 ISSP 2024: Digital Societies: Finnish Data")

digeff_plot + 
  theme(plot.margin = margin(10, 10, 10, 10))
```


## Institutional trust

```{r Inst trust,  fig.cap = "Distribution of institutional trust indicator", echo=FALSE}
# k48a plot
ggplot(AiR_clean, aes(x = k48a_f)) +
  geom_bar(fill="steelblue", color="black", alpha = 0.7) +
  labs(y = "Count",
       caption = "Data from FSD3994 ISSP 2024: Digital Societies: Finnish Data") +
  theme_pubr()

```



## Perceived personal threat of job loss

```{r Personal threat, fig.cap="Distribution of personal threat indicator"}
# k57 plot
ggplot(AiR_clean, aes(x = k57x_f)) +
  geom_bar(fill="steelblue", color="black", alpha = 0.7) +
  labs(y = "Count",
       caption = "Data from FSD3994 ISSP 2024: Digital Societies: Finnish Data") +
  scale_x_discrete(labels = get_labels(AiR_clean$k57x_f))+
  theme_pubr()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size=8))


```


\clearpage

# Demographics

## Gender

```{r k1 plot, fig.cap="Distribution of gender"}
gender_counts <- table(AiR_clean$k1)  
# convert to percentages
gender_percentages <- prop.table(gender_counts) * 100  

# create a data frame 
gender_df <- data.frame(
  Gender = names(gender_counts), 
  count = as.vector(gender_counts),  
  percentage = round(gender_percentages, 1)  
)

gender_df$Gender <- set_labels(gender_df$Gender, labels = gen3lik)

ggplot(gender_df, aes(x = Gender, y = count, fill=Gender)) +
  geom_bar(stat = "identity", alpha = 0.7, color = "black") +
  coord_flip() + 
  labs(y = "Count",
       caption = "Data from FSD3994 ISSP 2024: Digital Societies: Finnish Data") +
  scale_x_discrete(labels = get_labels(AiR_clean$k1_f))+
  geom_text(aes(label = paste0(percentage.Freq, "%")), 
            hjust = - 0.2, 
            color = "black") +
  scale_fill_manual(values = paletteer::paletteer_d("yarrr::basel", n = 3)) +
  theme_pubr()+
  theme(legend.position = "none")+
  scale_y_continuous(limits = c(0, 800))
```

## Age

```{r ika stats}
library(dplyr)

# Summary statistics for the variable "ika"
summary_ika <- AiR_clean %>%
  summarise(
    mean = mean(ika, na.rm = TRUE),
    sd = sd(ika, na.rm = TRUE),
    min = min(ika, na.rm = TRUE),
    Q1 = quantile(ika, 0.25, na.rm = TRUE),  
    Q3 = quantile(ika, 0.75, na.rm = TRUE),
    max = max(ika, na.rm = TRUE),
    na_count = sum(is.na(ika))
  )
```

```{r ika stat table, tab.cap="Summary statistics for age variable 'ika'"}

ika_table <- summary_ika %>%
  gt() %>%
  tab_header(
    title = "Age Distribution Summary"
  ) %>%
  cols_label(
    mean = "Mean",
    sd = "Standard Deviation",
    min = "Minimum",
    Q1 = "Quartile 1",
    Q3 = "Quartile 3",
    max = "Maximum",
    na_count = "NA Count"
  ) %>%
  fmt_number(
    columns = c(mean, sd, min, max),
    decimals = 2
  ) %>%
  fmt_number(
    columns = c(na_count),
    decimals = 0
  )
ika_table
```

```{r ika plot, fig.cap="Distribution of age by 5-year bracket"}
# plot for the variable "ika"
ggplot(AiR_clean, aes(x = ika, fill=k1_f)) +
  geom_histogram(binwidth = 5,color="black", alpha = 0.7) +
  labs(title = "Histogram of age distribution",
       y = "Count",
       caption = "Data from FSD3994 ISSP 2024: Digital Societies: Finnish Data") +
    scale_fill_manual(values = paletteer::paletteer_d("yarrr::basel", n = 3), 
                    labels = get_labels(AiR_clean$k1_f)) +  # Use factor labels for k1
  theme_pubr()

```

## Education level

```{r k4 plot, fig.cap="Distribution of education level"}
# plot for the variable k4
ggplot(AiR_clean, aes(x = k4_f, fill=k1_f)) +
  geom_bar(color="black", alpha = 0.7) +
  labs(title = "Bar plot of education level by gender",
       y = "Count",
       caption = "Data from FSD3994 ISSP 2024: Digital Societies: Finnish Data") +
    scale_x_discrete(labels = get_labels(AiR_clean$k4_f))+
    scale_fill_manual(values = paletteer::paletteer_d("yarrr::basel", n = 3), 
                    labels = get_labels(AiR_clean$k1_f)) + 
    theme_pubr()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size=7))


```




# Exploratory factor analysis

## EFA 1: Comfort with AI

```{r AI comfort corr full}
# Select the variables of interest

EFAAI <- subset(AiR_clean, select= c(k58a,k58b,k58c,k58d,k58e))

EFAAI <- EFAAI%>% rename(Medical = k58a,
                         Driving = k58b, 
                         Uncertainty = k58c,
                         Interaction = k58d,
                         Reliability = k58e)


# Calculate the correlation matrix
AI_mat <- cor(EFAAI, method = "spearman")


# plot the matrix
corrplot(AI_mat, method = "color", 
         type = "full",  
         order = "original", 
         tl.col = "black",
         tl.cex = .8,
         diag = TRUE,
         addgrid.col = "cadetblue",
         addrect = 1,
         number.font = 1,
         tl.srt = 45,
         addCoef.col = "white",  
         number.cex = 1,      
         title = "Correlation matrix of AI comfort", 
         mar=c(0,0,2,0))         

```

```{r AIR KMO}
# Digital literacy KMO test

KMO_EFAAI <- KMO(EFAAI)
print(KMO_EFAAI)

```

```{r AIR Bartlett, include=FALSE}

# Correlation matrix non-parametric monotonic ordinal factor
corr_mat_aiefa <- cor(EFAAI, method = "spearman")


# Bartlett test
bartlett_ai <- cortest.bartlett(corr_mat_aiefa, n = nrow(EFAAI))
print(bartlett_ai)

```

```{r AIR parallel analysis all, fig.cap="Parallel analysis test for AI Comfort (full sample)"}

# Parallel analysis test for number of factors
fa.parallel(EFAAI,
            fa = "fa",
            fm ="uls",
            main="",
            n.iter = 10,
            correct=.5, 
            sqrt=TRUE)

```

```{r AIR full EFA}

EFAAI <- subset(AiR_clean, select= c(k58a,k58b,k58c,k58d,k58e))


fa_ai <- fa(EFAAI, 
              nfactors = 2, 
              rotate = "Promax",
              cor = "spearman",
              fm ="uls", 
              scores = "regression")
print(fa_ai)
```

```{r AI EFA diag, fig.cap="AI comfort factor structure (full sample)"}
# Factor loading diagram digital literacy
colnames(fa_ai$loadings) <- c("A1", "A2")
fa.diagram(fa_ai, 
           marg=c(1,1,1,1),
           main="",
           l.cex = 0.8,
           e.size = .1,
           cut = 0.3)

```

```{r AI Factor loadings variable}
# AI Factor loading variable
fact_score_ai <- as.data.frame(fa_ai$scores)
fact_score_ai <- fact_score_ai %>% rename(A1 = ULS1)
fact_score_ai <- fact_score_ai %>% rename(A2 = ULS2)
fact_score_ai$AR1 <- set_label(fact_score_ai$AR1, label = "A1: AI Comfort (Communication)")
fact_score_ai$AR2 <- set_label(fact_score_ai$AR2, label = "A2: AI Comfort (Automation)")

```

```{r EFA A1 A2 loadings, fig.cap="Digital Efficacy factor analysis loadings"}
# Extract loadings and communalities
loadings1_ai <- fa_ai$loadings[, 1]  
loadings2_ai <- fa_ai$loadings[, 2]  
communalities_ai <- fa_ai$communality  

# Data frame for the table
df_ai <- data.frame(
  Indicator = c("Medical", "Driving", "Uncertainty", "Interaction", "Reliability"),
  Variable = rownames(fa_ai$loadings),  
  A1 = loadings1_ai,
  A2 = loadings2_ai,
  Communalities = communalities_ai          
)

# gt table
ai_table <- df_ai %>%
  gt() %>%
  tab_header(
    title = "AI Comfort EFA Results",
    subtitle = "Factor Loadings and Communalities"
  ) %>%
  cols_label(
    Indicator = "Comfort Measure",
    Variable = "Variable",
    A1 = "A1",
    A2 = "A2",
    Communalities = "h² Communalities"
  ) %>%
  fmt_number(
    columns = c(A1, A2, Communalities),
    decimals = 3  
  )
ai_table

```


```{r AR1 AR2 plot, fig.cap="Scatterplot demonstrating separation of factors"}
ggplot(
  data = fact_score_ai,
  mapping = aes(A2, y = A1)) +
  geom_point(color="steelblue4", alpha=0.5) +
  #geom_boxplot(fill="steelblue3", alpha=0.8, varwidth=TRUE, staplewidth = 0.2, coef=1.5) +
  labs(caption = "Data from FSD3994 ISSP 2024: Digital Societies: Finnish Data") +
  theme_pubr()
```


## EFA 2: Online Efficacy

```{r Correlation matrix, fig.cap="Correlation of digital efficacy variables"}
library(polycor)

# Select the variables of interest
Digeff_df <- AiR_clean[, c("k32ax", "k32bx", "k32cx", "k32dx")]
Digeff_df <- Digeff_df%>% rename(Usage = k32ax,
                                 Search = k32bx,
                                 Privacy = k32cx,
                                 Learning = k32dx)


# Calculate the polychoric correlation matrix
polychoric_correlation_matrix <- polychoric(Digeff_df)

# View the correlation matrix
Digeff_mat <-polychoric_correlation_matrix$rho

# plot the matrix
corrplot(Digeff_mat, method = "color", 
         type = "full",  
         order = "original", 
         tl.col = "black",
         tl.cex = .8,
         diag = TRUE,
         addgrid.col = "cadetblue",
         addrect = 1,
         number.font = 1,
         tl.srt = 45,
         addCoef.col = "white",  
         number.cex = 1,      
         title = "", 
         mar=c(0,0,2,0))         


```


```{r Digital efficacy check consistency.}
# Cronbach’s Alpha Digital Efficacy indicators
alpha_dl <- alpha(AiR_clean[, c("k32ax", "k32bx", "k32cx", "k32dx")])
print(alpha_dl$total$raw_alpha)

```

```{r DL KMO}
# Digital literacy KMO test

EFAAB <- subset(AiR_clean, select= c(k32ax,k32bx,k32cx,k32dx))
KMO_EFAAB <- KMO(EFAAB)
print(KMO_EFAAB)

```


```{r DL Bartlett}

# Bartlett test
bartlett_ab <- cortest.bartlett(Digeff_mat, n = nrow(EFAAB))
print(bartlett_ab)

```

```{r DL PA, fig.cap="Parallel analysis test for Digital Literacy factor extraction"}

# Parallel analysis test for number of factors
set.seed(123)
fa.parallel(EFAAB,
            fa = "fa",
            main="",
            cor = "poly",
            n.iter = 10,
            correct=.5, 
            sqrt=TRUE)

```


```{r DL EFA}

# EFA digital literacy
fa_dl <- fa(EFAAB, 
              nfactors = 1, 
              rotate = "Promax",
              cor = "poly",
              fm ="uls", 
              scores = "regression")
print(fa_dl)
```


```{r EFA D1 loadings, fig.cap="Digital Efficacy factor analysis loadings"}
# Extract loadings and communalities
loadings_d1 <- fa_dl$loadings[, 1]  
communalities_d1 <- fa_dl$communality  

# Data frame for the table
df_d1 <- data.frame(
  Indicator = c("Usage", "Search", "Privacy", "Learning"),
  Variable = rownames(fa_dl$loadings),  
  Loadings = loadings_d1,                  
  Communalities = communalities_d1          
)

# gt table
d1_table <- df_d1 %>%
  gt() %>%
  tab_header(
    title = "Digital Efficacy (D1) EFA Results",
    subtitle = "Factor Loadings and Communalities"
  ) %>%
  cols_label(
    Indicator = "Online Skill",
    Variable = "Variable",
    Loadings = "ULS Loadings",
    Communalities = "h² Communalities"
  ) %>%
  fmt_number(
    columns = c(Loadings, Communalities),
    decimals = 3  
  )
d1_table

```

```{r DL EFA diag, fig.cap="Digital Efficacy factor analysis diagram"}
# Factor loading diagram digital literacy
colnames(fa_dl$loadings) <- c("D1")
fa.diagram(fa_dl, 
           marg=c(1,1,1,1),
           main="",
           l.cex = 0.8,
           e.size = .1,
           cut = 0)

```

```{r DL Factor scores variable}
# DL Factor loading variable
fact_score_dl <- as.data.frame(fa_dl$scores)
fact_score_dl <- fact_score_dl %>% rename(D1 = ULS1)
fact_score_dl$D1 <- set_label(fact_score_dl$D1, label = "D1: Digital Efficacy")
summary(fact_score_dl$D1)
```

```{r merge factor scores to DF}
# Add extracted factors to data frame

AiRf <- cbind(AiR_clean, fact_score_dl, fact_score_ai)

```


# Latent Class Analysis

**Note:** Latent class are renumbered to maintain consistent alignment of distinct class profiles throughout the report. The ordering was a subjective choice.

```{r LCA class count selection}
# Select variables for class analysis
LCAAi <- subset(AiRf, select= c(k58a_f,k58b_f,k58c_f,k58d_f,k58e_f))
# set seed to reproduce results
set.seed(123)
# Define the formula for LCA
lca_ai_formula <- cbind(k58a_f, k58b_f, k58c_f, k58d_f, k58e_f) ~ 1  # No covariates

# Loop through different numbers of classes to evaluate fit
results_ai <- list()
for (n in 1:8) {
    results_ai[[n]] <- poLCA(lca_ai_formula, LCAAi, nclass = n, maxiter = 1000, na.rm = FALSE)
}
```

```{r AI LCA BIC AIC plot, fig.cap="BIC value plot"}
# Extract BIC/AIC values
bic_values <- sapply(results_ai, function(x) x$bic)
aic_values <- sapply(results_ai, function(x) x$aic)

# Plot BIC/AIC values
plot(1:8, bic_values, type = "b", col = "steelblue", xlab = "Number of Classes", ylab = "BIC", main = "BIC for Different Class Models")
points(1:8, aic_values, type = "b", col = "tomato")
legend("topright", legend = c("BIC", "AIC"), col = c("steelblue", "tomato"), pch = 1)
```

```{r AI recep LCA Results}
# set seed to reproduce results
set.seed(123)

# Fit the LCA model with a specified number of classes
lca_ai_model <- poLCA(lca_ai_formula, LCAAi, nclass = 4, maxiter = 1000, na.rm = FALSE)

# Manual Entropy Calculation for poLCA
error_prior <- lca_ai_model$posterior
n <- nrow(error_prior)
k <- ncol(error_prior)

# Calculate sum of posterior * log(posterior)
entropy_sum <- sum(-error_prior * log(error_prior), na.rm = TRUE)

# Relative Entropy formula
entropy <- 1 - (entropy_sum / (n * log(k)))
print(entropy)
```

```{r Summary of membership}
class_membership <- lca_ai_model$predclass
LCAAi$class <- class_membership
# Summarize class characteristics
table(LCAAi$class)
```

```{r Reorder latent classes and add to df}
LCAAi <- LCAAi %>%
  mutate(Class = case_when(
    class == 1 ~ 3,
    class == 2 ~ 1,
    class == 3 ~ 2,
    class == 4 ~ 4,
    TRUE ~ NA_real_  # In case there are any unexpected values
  ))

AiRf$Class <- LCAAi$Class

AiRf$Class_f <- as.factor(AiRf$Class)
AiRf$Class_f <- set_label(AiRf$Class_f, label = "Latent Class")
latclass4 <- c(
  `1` = "Adopters",
  `2` = "Cautious",
  `3` = "Resisters",
  `4` = "Rejecters"
  ) 
AiRf$Class_f <- set_labels(AiRf$Class_f, labels = latclass4)

```

```{r AI LCA analysis matrix, fig.cap="Classification table comparing LCA predicted vs. posterior probabilities"}
# Identify class with the highest posterior probability for each obs.
pred_class <- lca_ai_model$predclass

# compare predicted membership against the sum of posterior probabilities.
posterior_probs <- lca_ai_model$posterior

# Create the analysis matrix 
confusion_mat <- round(t(posterior_probs) %*% model.matrix(~ as.factor(pred_class) - 1), 2)
colnames(confusion_mat) <- paste(1:ncol(confusion_mat))
rownames(confusion_mat) <- paste(1:nrow(confusion_mat))

# Convert the confusion_mat to a data frame
conf_df <- melt(confusion_mat)

# Rename columns for clarity 
colnames(conf_df) <- c("Probabilistic_Class", "Assigned_Class", "Frequency")

# Renumber to align with report numbering
conf_df <- conf_df %>%
  mutate(Probabilistic = case_when(
    Probabilistic_Class == 1 ~ "Class 3",
    Probabilistic_Class == 2 ~ "Class 1",
    Probabilistic_Class == 3 ~ "Class 2",
    Probabilistic_Class == 4 ~ "Class 4"),
    Assigned = case_when(
    Assigned_Class == 1 ~ "Class 3",
    Assigned_Class == 2 ~ "Class 1",
    Assigned_Class == 3 ~ "Class 2",
    Assigned_Class == 4 ~ "Class 4"))

# Plot the categorization table (confusion matrix)
library(forcats)
ggplot(conf_df, aes(x = Assigned, y = fct_rev(Probabilistic), fill = Frequency)) +
  geom_tile(color = "white") + 
  geom_text(aes(label = Frequency), color = "black", size = 4) +
  scale_fill_gradient(low = "white", high = "steelblue") +
  labs(title = "Classification Diagnostic Matrix",
       x = "Assigned Class",
       y = "Probabilistic (Expected) Class",
       fill = "Frequency",
       caption = "Data from FSD3994 ISSP 2024: Digital Societies: Finnish Data") +
  coord_fixed()+
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size=8),
        axis.text.y = element_text(size=8), 
        legend.position = "none")

```


```{r Separate by class structures}
library(dplyr)

Adopters <- AiRf %>% filter(Class == 1)
Cautious <- AiRf %>% filter(Class == 2)
Resisters <- AiRf %>% filter(Class == 3)
Rejecters <- AiRf %>% filter(Class == 4)

```

# Class Analysis

## Class 1: "Adopters"

```{r class 1 ai attitudes, fig.cap="Class 1 AI receptivity indicator distributions", fig.height=6}
# k58a plot
k58a_plot1 <- ggplot(Adopters, aes(x = k58a_f)) +
  geom_bar(fill="#007A33FF", color="black", alpha = 0.7) +
  labs(y = "Count") +
  scale_x_discrete(limits = as.character(0:10)) +
  theme_pubr()
# k58b plot
k58b_plot1 <- ggplot(Adopters, aes(x = k58b_f)) +
  geom_bar(fill="#007A33FF", color="black", alpha = 0.7) +
  labs(y = "Count") +
  scale_x_discrete(limits = as.character(0:10)) +
  theme_pubr()
# k58c plot
k58c_plot1 <- ggplot(Adopters, aes(x = k58c_f)) +
  geom_bar(fill="#007A33FF", color="black", alpha = 0.7) +
  labs(y = "Count") +
  scale_x_discrete(limits = as.character(0:10)) +
  theme_pubr()
# k58d plot
k58d_plot1 <- ggplot(Adopters, aes(x = k58d_f)) +
  geom_bar(fill="#007A33FF", color="black", alpha = 0.7) +
  labs(y = "Count") +
  scale_x_discrete(limits = as.character(0:10)) +
  theme_pubr()
# k58e plot
k58e_plot1 <- ggplot(Adopters, aes(x = k58e_f)) +
  geom_bar(fill="#007A33FF", color="black", alpha = 0.7) +
  labs(y = "Count") +
  scale_x_discrete(limits = as.character(0:10)) +
  theme_pubr()


attitude_plot <- wrap_plots(
  k58a_plot1, k58b_plot1, 
  k58c_plot1, k58d_plot1, 
  k58e_plot1, 
  ncol = 2
) + 
  plot_annotation(caption = "Data from FSD3994 ISSP 2024: Digital Societies: Finnish Data")

attitude_plot + 
  theme(plot.margin = margin(10, 10, 10, 10))
```


## Class 2: "Cautious"

```{r class 2 ai attitudes, fig.cap="Class 2 AI receptivity indicator distributions", fig.height=6}
# k58a plot
k58a_plot2 <- ggplot(Cautious, aes(x = k58a_f)) +
  geom_bar(fill="#FE5000FF", color="black", alpha = 0.7) +
  labs(y = "Count") +
  scale_x_discrete(limits = as.character(0:10)) +
  theme_pubr()
# k58b plot
k58b_plot2 <- ggplot(Cautious, aes(x = k58b_f)) +
  geom_bar(fill="#FE5000FF", color="black", alpha = 0.7) +
  labs(y = "Count") +
  scale_x_discrete(limits = as.character(0:10)) +
  theme_pubr()
# k58c plot
k58c_plot2 <- ggplot(Cautious, aes(x = k58c_f)) +
  geom_bar(fill="#FE5000FF", color="black", alpha = 0.7) +
  labs(y = "Count") +
  scale_x_discrete(limits = as.character(0:10)) +
  theme_pubr()
# k58d plot
k58d_plot2 <- ggplot(Cautious, aes(x = k58d_f)) +
  geom_bar(fill="#FE5000FF", color="black", alpha = 0.7) +
  labs(y = "Count") +
  scale_x_discrete(limits = as.character(0:10)) +
  theme_pubr()
# k58e plot
k58e_plot2 <- ggplot(Cautious, aes(x = k58e_f)) +
  geom_bar(fill="#FE5000FF", color="black", alpha = 0.7) +
  labs(y = "Count") +
  scale_x_discrete(limits = as.character(0:10)) +
  theme_pubr()


attitude_plot <- wrap_plots(
  k58a_plot2, k58b_plot2, 
  k58c_plot2, k58d_plot2, 
  k58e_plot2, 
  ncol = 2
) + 
  plot_annotation(caption = "Data from FSD3994 ISSP 2024: Digital Societies: Finnish Data")

attitude_plot + 
  theme(plot.margin = margin(10, 10, 10, 10))
```


## Class 3: "Resisters"

```{r class 3 ai attitudes, fig.cap="Class 3 (Resisters) AI receptivity indicator distributions", fig.height=6}
# k58a plot
k58a_plot3 <- ggplot(Resisters, aes(x = k58a_f)) +
  geom_bar(fill="#0072CEFF", color="black", alpha = 0.7) +
  labs(y = "Count") +
  scale_x_discrete(limits = as.character(0:10)) +
  theme_pubr()
# k58b plot
k58b_plot3 <- ggplot(Resisters, aes(x = k58b_f)) +
  geom_bar(fill="#0072CEFF", color="black", alpha = 0.7) +
  labs(y = "Count") +
  scale_x_discrete(limits = as.character(0:10)) +
  theme_pubr()
# k58c plot
k58c_plot3 <- ggplot(Resisters, aes(x = k58c_f)) +
  geom_bar(fill="#0072CEFF", color="black", alpha = 0.7) +
  labs(y = "Count") +
  scale_x_discrete(limits = as.character(0:10)) +
  theme_pubr()
# k58d plot
k58d_plot3 <- ggplot(Resisters, aes(x = k58d_f)) +
  geom_bar(fill="#0072CEFF", color="black", alpha = 0.7) +
  labs(y = "Count") +
  scale_x_discrete(limits = as.character(0:10)) +
  theme_pubr()
# k58e plot
k58e_plot3 <- ggplot(Resisters, aes(x = k58e_f)) +
  geom_bar(fill="#0072CEFF", color="black", alpha = 0.7) +
  labs(y = "Count") +
  scale_x_discrete(limits = as.character(0:10)) +
  theme_pubr()


attitude_plot <- wrap_plots(
  k58a_plot3, k58b_plot3, 
  k58c_plot3, k58d_plot3, 
  k58e_plot3, 
  ncol = 2
) + 
  plot_annotation(caption = "Data from FSD3994 ISSP 2024: Digital Societies: Finnish Data")

attitude_plot + 
  theme(plot.margin = margin(10, 10, 10, 10))
```


## Class 4: "Rejecters"

```{r Rejecters ai attitudes, fig.cap="Class 4 (Rejecters) AI receptivity indicator distributions", fig.height=6}
# k58a plot
k58a_plot4 <- ggplot(Rejecters, aes(x = k58a_f)) +
  geom_bar(fill="#8D9093FF", color="black", alpha = 0.7) +
  labs(y = "Count") +
  scale_x_discrete(limits = as.character(0:10)) +
  theme_pubr()
# k58b plot
k58b_plot4 <- ggplot(Rejecters, aes(x = k58b_f)) +
  geom_bar(fill="#8D9093FF", color="black", alpha = 0.7) +
  labs(y = "Count") +
  scale_x_discrete(limits = as.character(0:10)) +
  theme_pubr()
# k58c plot
k58c_plot4 <- ggplot(Rejecters, aes(x = k58c_f)) +
  geom_bar(fill="#8D9093FF", color="black", alpha = 0.7) +
  labs(y = "Count") +
  scale_x_discrete(limits = as.character(0:10)) +
  theme_pubr()
# k58d plot
k58d_plot4 <- ggplot(Rejecters, aes(x = k58d_f)) +
  geom_bar(fill="#8D9093FF", color="black", alpha = 0.7) +
  labs(y = "Count") +
  scale_x_discrete(limits = as.character(0:10)) +
  theme_pubr()
# k58e plot
k58e_plot4 <- ggplot(Rejecters, aes(x = k58e_f)) +
  geom_bar(fill="#8D9093FF", color="black", alpha = 0.7) +
  labs(y = "Count") +
  scale_x_discrete(limits = as.character(0:10)) +
  theme_pubr()


attitude_plot <- wrap_plots(
  k58a_plot4, k58b_plot4, 
  k58c_plot4, k58d_plot4, 
  k58e_plot4, 
  ncol = 2
) + 
  plot_annotation(caption = "Data from FSD3994 ISSP 2024: Digital Societies: Finnish Data")

attitude_plot + 
  theme(plot.margin = margin(10, 10, 10, 10))
```

## Expected Values

```{r Expected values ai, fig.cap="AI omfort level expected values by latent class", fig.height=6}
# Create dataframes of expected values for each class
evi_df1 <- data.frame(Class = "Adopters",
                     Variable = 1:5,
                     Expected = c(mean(Adopters$k58a, na.rm = TRUE),
                                  mean(Adopters$k58b, na.rm = TRUE),
                                  mean(Adopters$k58c, na.rm = TRUE),
                                  mean(Adopters$k58d, na.rm = TRUE),
                                  mean(Adopters$k58e, na.rm = TRUE)))

evi_df2 <- data.frame(Class = "Cautious",
                     Variable = 1:5,
                     Expected = c(mean(Cautious$k58a, na.rm = TRUE),
                                  mean(Cautious$k58b, na.rm = TRUE),
                                  mean(Cautious$k58c, na.rm = TRUE),
                                  mean(Cautious$k58d, na.rm = TRUE),
                                  mean(Cautious$k58e, na.rm = TRUE)))

evi_df3 <- data.frame(Class = "Resisters",
                     Variable = 1:5,
                     Expected = c(mean(Resisters$k58a, na.rm = TRUE),
                                  mean(Resisters$k58b, na.rm = TRUE),
                                  mean(Resisters$k58c, na.rm = TRUE),
                                  mean(Resisters$k58d, na.rm = TRUE),
                                  mean(Resisters$k58e, na.rm = TRUE)))

evi_df4 <- data.frame(Class = "Rejecters",
                     Variable = 1:5,
                     Expected = c(mean(Rejecters$k58a, na.rm = TRUE),
                                  mean(Rejecters$k58b, na.rm = TRUE),
                                  mean(Rejecters$k58c, na.rm = TRUE),
                                  mean(Rejecters$k58d, na.rm = TRUE),
                                  mean(Rejecters$k58e, na.rm = TRUE)))

# Combine all data frames
evi_plot <- rbind(evi_df1, evi_df2, evi_df3, evi_df4)

# Define predictor variable labels
predi_vars <- c("Medical", "Driving", "Uncertainty", "Interaction", "Reliability")

# Convert Variables to factor with proper labels
evi_plot$Variable <- factor(evi_plot$Variable, levels = 1:5, labels = predi_vars)
evi_plot$Class <- factor(evi_plot$Class, levels = c("Adopters", "Cautious", "Resisters", "Rejecters"))

# Plot
ggplot(data = evi_plot, aes(x = Variable, y = Expected, color = Class, group = Class)) +
  geom_point(shape = 2, size = 2) +
  geom_line() +
  labs(x = "Outcome Variables", 
       y = "Expected Values", 
       color = "Construct",
       title = "Expected Values by Latent Class",
       subtitle = "Comfort level measured on 0-10 scale") +
  scale_color_manual(values = paletteer::paletteer_d("nbapalettes::knicks_holiday", n = 4))+
  theme_pubr() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size=8))+
  plot_annotation(caption = "Data from FSD3994 ISSP 2024: Digital Societies: Finnish Data")
```


# Latent Class Regression

**Note:** The responses were recoded from 0-10 to 1–11 to comply with the requirements of the poLCA package in R Studio but responses have been reported in terms of the survey values.


```{r LCR setup and model all}
library(dplyr)

# Prepare dataframe for poLCA regression
AiR_lcr <- AiRf %>% dplyr::select(k58a, k58b, k58c, k58d, k58e, D1, k32ax, k32bx, k32cx, k32dx, k48a, k57x, k1, ika, k4, Class, Class_f) 

# Recode outcome variables for poLCA regression
AiR_lcr$k58a <- as.numeric(as.character(AiR_lcr$k58a))
AiR_lcr$k58b <- as.numeric(as.character(AiR_lcr$k58b))
AiR_lcr$k58c <- as.numeric(as.character(AiR_lcr$k58c))
AiR_lcr$k58d <- as.numeric(as.character(AiR_lcr$k58d))
AiR_lcr$k58e <- as.numeric(as.character(AiR_lcr$k58e))
AiR_lcr$k58a1 <- AiR_lcr$k58a + 1
AiR_lcr$k58b1 <- AiR_lcr$k58b + 1
AiR_lcr$k58c1 <- AiR_lcr$k58c + 1
AiR_lcr$k58d1 <- AiR_lcr$k58d + 1
AiR_lcr$k58e1 <- AiR_lcr$k58e + 1

# Ensure Class is numeric
AiR_lcr$Class <- as.numeric(AiR_lcr$Class)

# Set Class_f as a factor with "1" as the reference level
AiR_lcr$Class_f <- relevel(as.factor(AiR_lcr$Class), ref = "1")

# Define the formula for the latent class regression
f_reg <- cbind(k58a1, k58b1, k58c1, k58d1, k58e1) ~ k1 + ika + k4 + D1 + k48a + k57x

# Perform latent class regression
set.seed(123)
LCA_reg <- poLCA(f_reg, data = AiR_lcr, nclass = 4, nrep = 10, verbose = TRUE)

```

```{r LCR setup and model}
library(dplyr)

# Prepare dataframe for poLCA regression

AiR_lcr <- AiRf %>% dplyr::select(k58a, k58b, k58c, k58d, k58e, D1, k48a, k57x, k1, ika, k4) 

# Recode outcome variables for poLCA regression
AiR_lcr$k58a <- as.numeric(as.character(AiR_lcr$k58a))
AiR_lcr$k58b <- as.numeric(as.character(AiR_lcr$k58b))
AiR_lcr$k58c <- as.numeric(as.character(AiR_lcr$k58c))
AiR_lcr$k58d <- as.numeric(as.character(AiR_lcr$k58d))
AiR_lcr$k58e <- as.numeric(as.character(AiR_lcr$k58e))
AiR_lcr$k58a1 <- as.factor(AiR_lcr$k58a + 1)
AiR_lcr$k58b1 <- as.factor(AiR_lcr$k58b + 1)
AiR_lcr$k58c1 <- as.factor(AiR_lcr$k58c + 1)
AiR_lcr$k58d1 <- as.factor(AiR_lcr$k58d + 1)
AiR_lcr$k58e1 <- as.factor(AiR_lcr$k58e + 1)



# Define the formula for the latent class regression
f_reg <- cbind(k58a1, k58b1, k58c1, k58d1, k58e1) ~ k1 + ika + k4 + D1 + k48a + k57x

# Perform latent class regression
set.seed(123)
LCA_reg <- poLCA(f_reg, data = AiR_lcr, nclass = 4, nrep = 10, verbose = TRUE)

```

```{r LCR retrieve data}
# Extract coefficients and SEs
coeffs <- LCA_reg$coeff
ses <- LCA_reg$coeff.se

# Calculate t-values and p-values
t_values <- coeffs / ses
p_values <- 2 * (1 - pnorm(abs(t_values)))

# Format into a data frame
lcr_df <- data.frame(
  Variable = rep(rownames(coeffs), ncol(coeffs)),
  Class = rep(2:(ncol(coeffs)+1), each = nrow(coeffs)), # Class 1 is reference
  Coeff = round(as.vector(coeffs),3),
  Coeff.SE = round(as.vector(ses),3),
  P_Value = round(as.vector(p_values),3)
)

# Add significance stars for readability 
lcr_df$Significance <- cut(lcr_df$P_Value, 
                           breaks = c(-Inf, 0.001, 0.01, 0.05, 0.1, Inf), 
                           labels = c("***", "**", "*", ".", " "))

# View the result
lcr_df$Class <- as.factor(lcr_df$Class)

# Match classes to profiles based on conditional probs
lcrclass3 <- c(
  `2` = "Cautious",
  `3` = "Resisters",
  `4` = "Rejecters"
  ) 

lcr_df$Class <- factor(lcr_df$Class, levels = c("2", "4", "3"))
lcr_df$Class <- set_labels(lcr_df$Class, labels = lcrclass3)
lcr_df <- lcr_df %>%
  mutate(
    lower_ci = Coeff - 1.96 * Coeff.SE,
    upper_ci = Coeff + 1.96 * Coeff.SE
  )
# Calculate confidence interval limits
lcr_df$lower_ci <- round(lcr_df$lower_ci,3)
lcr_df$upper_ci <- round(lcr_df$upper_ci,3)

lcrc_plot <- lcr_df %>%
  filter(Variable != "(Intercept)")

```

```{r LCR coefficient labels}

lcrvar7 <- c(
  k1 = "Cautious",
  ika = "Resisters",
  k4 = "Rejecters",
  D1 = "Digital Efficacy",
  k48a = "Trust parliament",
  k57x = "Perceive threat",
  `(Intercept)` = "Intercept"
  )
lcr_df$Variable <- set_labels(lcr_df$Variable, labels = lcrvar7)

lcrvar6 <- c(
  D1 = "Digital Efficacy",
  ika = "Age",
  k1 = "Gender",
  k4 = "Education",
  k48a = "Trust parliament",
  k57x = "Perceive threat")
lcrc_plot$Variable <- set_labels(lcrc_plot$Variable, labels = lcrvar6)

```


```{r LCR coefficient table, tab.cap="Latent class regression coefficients"}

# Create a summary variable column
lcr_df$Coeff_all <- paste(lcr_df$Coeff, 
                                 lcr_df$Significance, 
                                 " (", 
                                 lcr_df$Coeff.SE, 
                                 ")")
# Create columns of summary variable by class
lcr_dat <- lcr_df %>% dplyr::select(Class, Variable, Coeff_all)
lcr_dat_w <- lcr_dat %>% pivot_wider(names_from = Class, 
              values_from = Coeff_all)
# Rename classes for consistency
lcr_dat_w <- lcr_dat_w %>% rename("Class2"= `2`,
                              "Class3"= `4`,
                              "Class4"= `3`)

# Add column for indicator names
Indicator <- c("Constant", "Gender", "Age", "Education level", "Digital efficacy", "Institutional trust", "Perceived threat")
lcr_tab <- cbind(lcr_dat_w,Indicator)
# Reorder the table for readability
lcr_table <- lcr_tab %>% dplyr::select(Indicator, Variable, Class2, Class3, Class4)

# Create report table
LCR_table <- lcr_table %>%
  gt() %>%
  tab_header(
    title = "Latent Class Regression",
    subtitle = "Coeffients, significance code, and (Standard Error)"
  ) %>%
  cols_label(
    Class2 = "Class 2",
    Class3 = "Class 3",
    Class4 = "Class 4"
  ) %>% 
  tab_footnote(footnote = "Note: . p<0.1; *p<0.05; **p<0.01; ***p<0.001")

LCR_table

```

```{r LCR class2, tab.cap="Class 2 log-odds coefficients"}
# Create report table
lcr_class2 <- lcr_df %>% dplyr::filter(Class==2) %>% dplyr::select(Variable,Coeff,Coeff.SE,P_Value,Significance)

LCR_table2 <- lcr_class2 %>%
  gt() %>%
  tab_header(
    subtitle = "Class 1 to Class 2",
    title= "Log-odds coefficients"
  ) %>%
  cols_label(
    Variable = "Variable",
    Coeff = "Coefficient",
    Coeff.SE = "Standard Error",
    P_Value = "P-value"
  ) %>%
  tab_footnote(footnote = "Note: . p<0.1; *p<0.05; **p<0.01; ***p<0.001")

LCR_table2

```

```{r LCR class3, tab.cap="Class 3 log-odds coefficients"}
# Create report table
lcr_class3 <- lcr_df %>% dplyr::filter(Class==4) %>% dplyr::select(Variable,Coeff,Coeff.SE,P_Value,Significance)

LCR_table3 <- lcr_class3 %>%
  gt() %>%
  tab_header(
    subtitle = "Class 1 to Class 3",
    title= "Log-odds coefficients"
  ) %>%
  cols_label(
    Variable = "Variable",
    Coeff = "Coefficient",
    Coeff.SE = "Standard Error",
    P_Value = "P-value"
  ) %>%
  tab_footnote(footnote = "Note: . p<0.1; *p<0.05; **p<0.01; ***p<0.001")

LCR_table3

```

```{r LCR class4, tab.cap="Class 4 log-odds coefficients"}
# Create report table
lcr_class4 <- lcr_df %>% dplyr::filter(Class==3) %>% dplyr::select(Variable,Coeff,Coeff.SE,P_Value,Significance)

LCR_table4 <- lcr_class4 %>%
  gt() %>%
  tab_header(
    subtitle = "Class 1 to Class 4",
    title= "Log-odds coefficients"
  ) %>%
  cols_label(
    Variable = "Variable",
    Coeff = "Coefficient",
    Coeff.SE = "Standard Error",
    P_Value = "P-value"
  ) %>%
  tab_footnote(footnote = "Note: . p<0.1; *p<0.05; **p<0.01; ***p<0.001")

LCR_table4

```

**Note:** Confidence intervals were manually calculated as Coefficient plus or minus 1.96 x Coefficient SE.

```{r LCR coeff plot1, fig.cap="Plot of latent class regression coefficients"}

coeff_plot1 <- ggplot(lcr_df, aes(x = Coeff, y = Variable, color = Class)) +
  geom_point(size = 3) +
  geom_vline(xintercept = 0, linetype = 2)+
  geom_errorbar(aes(xmin = lower_ci, xmax = upper_ci), width = 0.2) +
  labs(title = "Latent Class Regression Coefficients",
       x = "Coefficient Estimate",
       y = "Predictor Variables") +
  theme_pubr() +
  scale_color_manual(values = paletteer::paletteer_d("nbapalettes::knicks_holiday")[c(2, 3, 4)], 
                    labels = get_labels(lcr_df$Class)) 

```

```{r LCR coeff plot2, fig.cap="Plot of latent class regression coefficients (coefficients only)"}

lcrc_plot$Variable <- as.factor(lcrc_plot$Variable)
coeff_plot2 <- ggplot(lcrc_plot, aes(y = Coeff, x = Variable, color = Class)) +
  geom_point(size = 3) +
  geom_hline(yintercept = 0, linetype = 2)+
  geom_errorbar(aes(ymin = lower_ci, ymax = upper_ci), width = 0.2) +
  labs(subtitle = "Predictor coefficients only:",
       y = "Coefficient Estimate",
       x = "Predictor Variables") +
  theme_pubr() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1,size = 8))+
  scale_x_discrete(labels = c(
    D1 = "D1 (Digital Efficacy)",
    ika = "ika (Age)",
    k1 = "k1 (Gender)",
    k4 = "k4 (Education)",
    k48a = "k48a (Trust parliament)",
    k57x = "k57x (Perceive threat)"))+
  scale_color_manual(values = paletteer::paletteer_d("nbapalettes::knicks_holiday")[c(2, 3, 4)], 
                    labels = get_labels(lcr_df$Class)) 


```

```{r LCR coeff plot group, fig.cap="Plot of latent class regression coefficients", fig.height=10, fig.width=7}

coeff_plots <- wrap_plots(
  coeff_plot1, coeff_plot2, 
  ncol = 1
) + 
  plot_annotation(caption = "Data from FSD3994 ISSP 2024: Digital Societies: Finnish Data")

coeff_plots + 
  theme(plot.margin = margin(10, 10, 10, 10))
```


```{r class 1 predictor plots}

# k32a plot
k32a_class1 <- ggplot(Adopters, aes(x = k32ax_f)) +
  geom_bar(fill="#007A33FF", color="black", alpha = 0.7) +
  labs(y = "Count") +
  theme_pubr()
# k32b plot
k32b_class1 <- ggplot(Adopters, aes(x = k32bx_f)) +
  geom_bar(fill="#007A33FF", color="black", alpha = 0.7) +
  labs(y = "Count") +
  theme_pubr()
# k32c plot
k32c_class1 <- ggplot(Adopters, aes(x = k32cx_f)) +
  geom_bar(fill="#007A33FF", color="black", alpha = 0.7) +
  labs(y = "Count") +
  theme_pubr()
# k32d plot
k32d_class1 <- ggplot(Adopters, aes(x = k32dx_f)) +
  geom_bar(fill="#007A33FF", color="black", alpha = 0.7) +
  labs(y = "Count") +
  theme_pubr()
k48a_class1 <- ggplot(Adopters, aes(x = k48a_f)) +
  geom_bar(fill="#007A33FF", color="black", alpha = 0.7) +
  labs(subtitle = "Adopters", y = "Count") +
  theme_pubr()
k57x_class1 <- ggplot(Adopters, aes(x = k57x_f)) +
  geom_bar(fill="#007A33FF", color="black", alpha = 0.7) +
  labs(subtitle = "Adopters", y = "Count") +
  theme_pubr()
```

```{r class 2 pred plots}

# k32a plot
k32a_class2 <- ggplot(Cautious, aes(x = k32ax_f)) +
  geom_bar(fill="#FE5000FF", color="black", alpha = 0.7) +
  labs(y = "Count") +
  theme_pubr()
# k32b plot
k32b_class2 <- ggplot(Cautious, aes(x = k32bx_f)) +
  geom_bar(fill="#FE5000FF", color="black", alpha = 0.7) +
  labs(y = "Count") +
  theme_pubr()
# k32c plot
k32c_class2 <- ggplot(Cautious, aes(x = k32cx_f)) +
  geom_bar(fill="#FE5000FF", color="black", alpha = 0.7) +
  labs(y = "Count") +
  theme_pubr()
# k32d plot
k32d_class2 <- ggplot(Cautious, aes(x = k32dx_f)) +
  geom_bar(fill="#FE5000FF", color="black", alpha = 0.7) +
  labs(y = "Count") +
  theme_pubr()
k48a_class2 <- ggplot(Cautious, aes(x = k48a_f)) +
  geom_bar(fill="#FE5000FF", color="black", alpha = 0.7) +
  labs(subtitle = "Cautious", y = "Count") +
  theme_pubr()
k57x_class2 <- ggplot(Cautious, aes(x = k57x_f)) +
  geom_bar(fill="#FE5000FF", color="black", alpha = 0.7) +
  labs(subtitle = "Cautious", y = "Count") +
  theme_pubr()
```

```{r class 3 pred plots}

# k32a plot
k32a_class3 <- ggplot(Resisters, aes(x = k32ax_f)) +
  geom_bar(fill="#0072CEFF", color="black", alpha = 0.7) +
  labs(y = "Count") +
  theme_pubr()
# k32b plot
k32b_class3 <- ggplot(Resisters, aes(x = k32bx_f)) +
  geom_bar(fill="#0072CEFF", color="black", alpha = 0.7) +
  labs(y = "Count") +
  theme_pubr()
# k32c plot
k32c_class3 <- ggplot(Resisters, aes(x = k32cx_f)) +
  geom_bar(fill="#0072CEFF", color="black", alpha = 0.7) +
  labs(y = "Count") +
  theme_pubr()
# k32d plot
k32d_class3 <- ggplot(Resisters, aes(x = k32dx_f)) +
  geom_bar(fill="#0072CEFF", color="black", alpha = 0.7) +
  labs(y = "Count") +
  theme_pubr()
k48a_class3 <- ggplot(Resisters, aes(x = k48a_f)) +
  geom_bar(fill="#0072CEFF", color="black", alpha = 0.7) +
  labs(subtitle = "Resisters",y = "Count") +
  theme_pubr()
k57x_class3 <- ggplot(Resisters, aes(x = k57x_f)) +
  geom_bar(fill="#0072CEFF", color="black", alpha = 0.7) +
  labs(subtitle = "Resisters", y = "Count") +
  theme_pubr()
```

```{r class 4 pred plots}

# k32a plot
k32a_class4 <- ggplot(Rejecters, aes(x = k32ax_f)) +
  geom_bar(fill="#8D9093FF", color="black", alpha = 0.7) +
  labs(y = "Count") +
  theme_pubr()
# k32b plot
k32b_class4 <- ggplot(Rejecters, aes(x = k32bx_f)) +
  geom_bar(fill="#8D9093FF", color="black", alpha = 0.7) +
  labs(y = "Count") +
  theme_pubr()
# k32c plot
k32c_class4 <- ggplot(Rejecters, aes(x = k32cx_f)) +
  geom_bar(fill="#8D9093FF", color="black", alpha = 0.7) +
  labs(y = "Count") +
  theme_pubr()
# k32d plot
k32d_class4 <- ggplot(Rejecters, aes(x = k32dx_f)) +
  geom_bar(fill="#8D9093FF", color="black", alpha = 0.7) +
  labs(y = "Count") +
  theme_pubr()
k48a_class4 <- ggplot(Rejecters, aes(x = k48a_f)) +
  geom_bar(fill="#8D9093FF", color="black", alpha = 0.7) +
  labs(subtitle = "Rejecters", y = "Count") +
  theme_pubr()
k57x_class4 <- ggplot(Rejecters, aes(x = k57x_f)) +
  geom_bar(fill="#8D9093FF", color="black", alpha = 0.7) +
  labs(subtitle = "Rejecters", y = "Count") +
  theme_pubr()
```

```{r class 1 demo plots}

gender_counts_class1 <- table(Adopters$k1)  
# convert to percentages
gender_percentages_class1 <- prop.table(gender_counts_class1) * 100  

# create a data frame 
gender_df_class1 <- data.frame(
  Gender = names(gender_counts_class1), 
  count = as.vector(gender_counts_class1),  
  percentage = round(gender_percentages_class1, 1)  
)

gender_df_class1$Gender <- set_labels(gender_df_class1$Gender, labels = gen3lik)

gender_class1 <- ggplot(gender_df_class1, aes(x = Gender, y = count, fill=Gender)) +
  geom_bar(stat = "identity", alpha = 0.7, color = "black") +
  coord_flip() + 
  labs(subtitle = "Adopters",y = "Count") +
  geom_text(aes(label = paste0(percentage.Freq, "%")), 
            hjust = - 0.2, 
            color = "black") +
  scale_fill_manual(values = paletteer::paletteer_d("yarrr::basel", n = 3)) +
  theme_pubr()+
  theme(legend.position = "none")+
  scale_y_continuous(limits = c(0, 300))

# plot for the variable "ika"
age_class1 <- ggplot(Adopters, aes(x = ika, fill=k1_f)) +
  geom_histogram(binwidth = 5,color="black", alpha = 0.7) +
  labs(subtitle = "Adopters",y = "Count") +
    scale_fill_manual(values = paletteer::paletteer_d("yarrr::basel", n = 3)) +  
  theme_pubr() +
  theme(legend.position = "none")

# plot for the variable k4
edu_class1 <- ggplot(Adopters, aes(x = k4_f, fill=k1_f)) +
  geom_bar(color="black", alpha = 0.7) +
  labs(subtitle = "Adopters",y = "Count") +
    scale_fill_manual(values = paletteer::paletteer_d("yarrr::basel", n = 3), 
                    labels = get_labels(AiR$k1_f)) + 
  theme_pubr() +
  theme(legend.position = "bottom")

```

```{r class 2 demo plots}

gender_counts_class2 <- table(Cautious$k1)  
# convert to percentages
gender_percentages_class2 <- prop.table(gender_counts_class2) * 100  

# create a data frame 
gender_df_class2 <- data.frame(
  Gender = names(gender_counts_class2), 
  count = as.vector(gender_counts_class2),  
  percentage = round(gender_percentages_class2, 1)  
)

gender_df_class2$Gender <- set_labels(gender_df_class2$Gender, labels = gen3lik)

gender_class2 <- ggplot(gender_df_class2, aes(x = Gender, y = count, fill=Gender)) +
  geom_bar(stat = "identity", alpha = 0.7, color = "black") +
  coord_flip() + 
  labs(subtitle = "Cautious",y = "Count") +
  geom_text(aes(label = paste0(percentage.Freq, "%")), 
            hjust = - 0.2, 
            color = "black") +
  scale_fill_manual(values = paletteer::paletteer_d("yarrr::basel", n = 3)) +
  theme_pubr()+
  theme(legend.position = "none")+
  scale_y_continuous(limits = c(0, 300))

# plot for the variable "ika"
age_class2 <- ggplot(Cautious, aes(x = ika, fill=k1_f)) +
  geom_histogram(binwidth = 5,color="black", alpha = 0.7) +
  labs(subtitle = "Cautious",y = "Count") +
    scale_fill_manual(values = paletteer::paletteer_d("yarrr::basel", n = 3)) +  
  theme_pubr() +
  theme(legend.position = "none")

# plot for the variable k4
edu_class2 <- ggplot(Cautious, aes(x = k4_f, fill=k1_f)) +
  geom_bar(color="black", alpha = 0.7) +
  labs(subtitle = "Cautious", y = "Count") +
    scale_fill_manual(values = paletteer::paletteer_d("yarrr::basel", n = 3), 
                    labels = get_labels(AiR$k1_f)) + 
  theme_pubr() +
  theme(legend.position = "bottom")

```

```{r class 3 demo plots}

gender_counts_class3 <- table(Resisters$k1)  
# convert to percentages
gender_percentages_class3 <- prop.table(gender_counts_class3) * 100  

# create a data frame 
gender_df_class3 <- data.frame(
  Gender = names(gender_counts_class3), 
  count = as.vector(gender_counts_class3),  
  percentage = round(gender_percentages_class3, 1)  
)

gender_df_class3$Gender <- set_labels(gender_df_class3$Gender, labels = gen3lik)

gender_class3 <- ggplot(gender_df_class3, aes(x = Gender, y = count, fill=Gender)) +
  geom_bar(stat = "identity", alpha = 0.7, color = "black") +
  coord_flip() + 
  labs(subtitle = "Resisters", y = "Count") +
  geom_text(aes(label = paste0(percentage.Freq, "%")), 
            hjust = - 0.2, 
            color = "black") +
  scale_fill_manual(values = paletteer::paletteer_d("yarrr::basel", n = 3)) +
  theme_pubr()+
  theme(legend.position = "none")+
  scale_y_continuous(limits = c(0, 300))

# plot for the variable "ika"
age_class3 <- ggplot(Resisters, aes(x = ika, fill=k1_f)) +
  geom_histogram(binwidth = 5,color="black", alpha = 0.7) +
  labs(subtitle = "Resisters", y = "Count") +
    scale_fill_manual(values = paletteer::paletteer_d("yarrr::basel", n = 3)) +  
  theme_pubr() +
  theme(legend.position = "none")

# plot for the variable k4
edu_class3 <- ggplot(Resisters, aes(x = k4_f, fill=k1_f)) +
  geom_bar(color="black", alpha = 0.7) +
  labs(subtitle = "Resisters", y = "Count") +
    scale_fill_manual(values = paletteer::paletteer_d("yarrr::basel", n = 3), 
                    labels = get_labels(AiR$k1_f)) + 
  theme_pubr() +
  theme(legend.position = "bottom")

```

```{r class 4 demo plots}

gender_counts_class4 <- table(Rejecters$k1)  
# convert to percentages
gender_percentages_class4 <- prop.table(gender_counts_class4) * 100  

# create a data frame 
gender_df_class4 <- data.frame(
  Gender = names(gender_counts_class4), 
  count = as.vector(gender_counts_class4),  
  percentage = round(gender_percentages_class4, 1)  
)

gender_df_class4$Gender <- set_labels(gender_df_class4$Gender, labels = gen3lik)

gender_class4 <- ggplot(gender_df_class4, aes(x = Gender, y = count, fill=Gender)) +
  geom_bar(stat = "identity", alpha = 0.7, color = "black") +
  coord_flip() + 
  labs(subtitle = "Rejecters",y = "Count") +
  geom_text(aes(label = paste0(percentage.Freq, "%")), 
            hjust = - 0.2, 
            color = "black") +
  scale_fill_manual(values = paletteer::paletteer_d("yarrr::basel", n = 3)) +
  theme_pubr()+
  theme(legend.position = "none")+
  scale_y_continuous(limits = c(0, 300))

# plot for the variable "ika"
age_class4 <- ggplot(Rejecters, aes(x = ika, fill=k1_f)) +
  geom_histogram(binwidth = 5,color="black", alpha = 0.7) +
  labs(subtitle = "Rejecters", 
       y = "Count") +
    scale_fill_manual(values = paletteer::paletteer_d("yarrr::basel", n = 3)) +  
  theme_pubr() +
  theme(legend.position = "none")

# plot for the variable k4
edu_class4 <- ggplot(Rejecters, aes(x = k4_f, fill=k1_f)) +
  geom_bar(color="black", alpha = 0.7) +
  labs(subtitle = "Rejecters", y = "Count") +
    scale_fill_manual(values = paletteer::paletteer_d("yarrr::basel", n = 3), 
                    labels = get_labels(AiR$k1_f)) + 
  theme_pubr() +
  theme(legend.position = "bottom")

```

```{r class 1 ika stats, tab.cap="Summary statistics for age variable 'ika'"}
library(dplyr)

# Summary statistics for the variable "ika"
summary_ika_class1 <- Adopters %>%
  summarise(
    mean = mean(ika, na.rm = TRUE),
    sd = sd(ika, na.rm = TRUE),
    min = min(ika, na.rm = TRUE),
    Q1 = quantile(ika, 0.25, na.rm = TRUE),  
    Q3 = quantile(ika, 0.75, na.rm = TRUE),
    max = max(ika, na.rm = TRUE),
    na_count = sum(is.na(ika))
  )

ika_table_class1 <- summary_ika_class1 %>%
  gt() %>%
  tab_header(
    title = "Class 1 (Adopters) age summary statistics"
  ) %>%
  cols_label(
    mean = "Mean",
    sd = "Standard Deviation",
    min = "Minimum",
    Q1 = "Quartile 1",
    Q3 = "Quartile 3",
    max = "Maximum",
    na_count = "NA Count"
  ) %>%
  fmt_number(
    columns = c(mean, sd, min, max),
    decimals = 2
  ) %>%
  fmt_number(
    columns = c(na_count),
    decimals = 0
  )
ika_table_class1
```

```{r class 2 ika stats, tab.cap="Summary statistics for age variable 'ika'"}
library(dplyr)

# Summary statistics for the variable "ika"
summary_ika_class2 <- Cautious %>%
  summarise(
    mean = mean(ika, na.rm = TRUE),
    sd = sd(ika, na.rm = TRUE),
    min = min(ika, na.rm = TRUE),
    Q1 = quantile(ika, 0.25, na.rm = TRUE),  
    Q3 = quantile(ika, 0.75, na.rm = TRUE),
    max = max(ika, na.rm = TRUE),
    na_count = sum(is.na(ika))
  )

ika_table_class2 <- summary_ika_class2 %>%
  gt() %>%
  tab_header(
    title = "Class 2 (Cautious) age summary statistics"
  ) %>%
  cols_label(
    mean = "Mean",
    sd = "Standard Deviation",
    min = "Minimum",
    Q1 = "Quartile 1",
    Q3 = "Quartile 3",
    max = "Maximum",
    na_count = "NA Count"
  ) %>%
  fmt_number(
    columns = c(mean, sd, min, max),
    decimals = 2
  ) %>%
  fmt_number(
    columns = c(na_count),
    decimals = 0
  )
ika_table_class2
```

```{r class 3 ika stats, tab.cap="Summary statistics for age variable 'ika'"}
library(dplyr)

# Summary statistics for the variable "ika"
summary_ika_class3 <- Resisters %>%
  summarise(
    mean = mean(ika, na.rm = TRUE),
    sd = sd(ika, na.rm = TRUE),
    min = min(ika, na.rm = TRUE),
    Q1 = quantile(ika, 0.25, na.rm = TRUE),  
    Q3 = quantile(ika, 0.75, na.rm = TRUE),
    max = max(ika, na.rm = TRUE),
    na_count = sum(is.na(ika))
  )

ika_table_class3 <- summary_ika_class3 %>%
  gt() %>%
  tab_header(
    title = "Class 3 (Resisters) age summary statistics"
  ) %>%
  cols_label(
    mean = "Mean",
    sd = "Standard Deviation",
    min = "Minimum",
    Q1 = "Quartile 1",
    Q3 = "Quartile 3",
    max = "Maximum",
    na_count = "NA Count"
  ) %>%
  fmt_number(
    columns = c(mean, sd, min, max),
    decimals = 2
  ) %>%
  fmt_number(
    columns = c(na_count),
    decimals = 0
  )
ika_table_class3
```

```{r class 4 ika stats}
library(dplyr)

# Summary statistics for the variable "ika"
summary_ika_class4 <- Rejecters %>%
  summarise(
    mean = mean(ika, na.rm = TRUE),
    sd = sd(ika, na.rm = TRUE),
    min = min(ika, na.rm = TRUE),
    Q1 = quantile(ika, 0.25, na.rm = TRUE),  
    Q3 = quantile(ika, 0.75, na.rm = TRUE),
    max = max(ika, na.rm = TRUE),
    na_count = sum(is.na(ika))
  )

ika_table_class4 <- summary_ika_class4 %>%
  gt() %>%
  tab_header(
    title = "Class 4 (Rejecters) age summary statistics"
  ) %>%
  cols_label(
    mean = "Mean",
    sd = "Standard Deviation",
    min = "Minimum",
    Q1 = "Quartile 1",
    Q3 = "Quartile 3",
    max = "Maximum",
    na_count = "NA Count"
  ) %>%
  fmt_number(
    columns = c(mean, sd, min, max),
    decimals = 2
  ) %>%
  fmt_number(
    columns = c(na_count),
    decimals = 0
  )
ika_table_class4
```

## Digital Efficacy by Latent Class

```{r digeff plots class, fig.cap="Digital efficacy factor by latent class", fig.height=6}

ggplot(AiRf, aes(x=Class_f, y = D1, fill = Class_f)) +
  geom_boxplot(color="black", alpha = 0.7) +
  labs(x="Construct") +
  theme_pubr()+
  scale_fill_manual(values = paletteer::paletteer_d("nbapalettes::knicks_holiday", n = 4))+
  scale_x_discrete(labels = get_labels(AiRf$Class_f))+
  plot_annotation(caption = "Data from FSD3994 ISSP 2024: Digital Societies: Finnish Data") +
  theme(legend.position = "none")
```


## Institutional Trust by Latent Class

```{r inst trust by class plots, fig.cap="Age distribution by latent class", fig.height=6}

trust_class <- wrap_plots(
  k48a_class1, k48a_class2, 
  k48a_class3, k48a_class4,
  ncol = 2
) + 
  plot_annotation(caption = "Data from FSD3994 ISSP 2024: Digital Societies: Finnish Data")

trust_class + 
  theme(plot.margin = margin(10, 10, 10, 10))


```

## Perceived Threat by Latent Class

```{r perc threat by class plots, fig.cap="Perceived threat distribution by latent class", fig.height=6}

threat_class <- wrap_plots(
  k57x_class1, k57x_class2, 
  k57x_class3, k57x_class4,
  ncol = 2
) + 
  plot_annotation(caption = "Data from FSD3994 ISSP 2024: Digital Societies: Finnish Data")

threat_class + 
  theme(plot.margin = margin(10, 10, 10, 10))


```

## Gender by Latent Class

```{r gender by class plots, fig.cap="Gender distribution by latent class", fig.height=6}

gender_class <- wrap_plots(
  gender_class1, gender_class2, 
  gender_class3, gender_class4,
  ncol = 2
) + 
  plot_annotation(caption = "Data from FSD3994 ISSP 2024: Digital Societies: Finnish Data")

gender_class + 
  theme(plot.margin = margin(10, 10, 10, 10))


```

## Age by Latent Class

```{r age by class plots, fig.cap="Age distribution by latent class", fig.height=6}

age_class <- wrap_plots(
  age_class1, age_class2, 
  age_class3, age_class4,
  ncol = 2
) + 
  plot_annotation(caption = "Data from FSD3994 ISSP 2024: Digital Societies: Finnish Data")

age_class + 
  theme(plot.margin = margin(10, 10, 10, 10))


```

## Education level by latent class

```{r edu by class plots, fig.cap="Education level distribution by latent class", fig.height=6}

edu_class <- wrap_plots(
  edu_class1, edu_class2, 
  edu_class3, edu_class4,
  ncol = 2
) + 
  plot_annotation(caption = "Data from FSD3994 ISSP 2024: Digital Societies: Finnish Data")

edu_class + 
  theme(plot.margin = margin(10, 10, 10, 10))


```

## Expected Values of predictors

```{r Expected values pred, fig.cap="Predictor expected values by latent class", fig.height=6}
# Create dataframes of expected values for each class
ev_df1 <- data.frame(Class = "Adopters",
                     Variable = 1:6,
                     Expected = c(mean(Adopters$k32ax, na.rm = TRUE),
                                  mean(Adopters$k32bx, na.rm = TRUE),
                                  mean(Adopters$k32cx, na.rm = TRUE),
                                  mean(Adopters$k32dx, na.rm = TRUE),
                                  0.5 * mean(Adopters$k48a, na.rm = TRUE),
                                  mean(Adopters$k57x, na.rm = TRUE)))

ev_df2 <- data.frame(Class = "Cautious",
                     Variable = 1:6,
                     Expected = c(mean(Cautious$k32ax, na.rm = TRUE),
                                  mean(Cautious$k32bx, na.rm = TRUE),
                                  mean(Cautious$k32cx, na.rm = TRUE),
                                  mean(Cautious$k32dx, na.rm = TRUE),
                                  0.5 * mean(Cautious$k48a, na.rm = TRUE),
                                  mean(Cautious$k57x, na.rm = TRUE)))

ev_df3 <- data.frame(Class = "Resisters",
                     Variable = 1:6,
                     Expected = c(mean(Resisters$k32ax, na.rm = TRUE),
                                  mean(Resisters$k32bx, na.rm = TRUE),
                                  mean(Resisters$k32cx, na.rm = TRUE),
                                  mean(Resisters$k32dx, na.rm = TRUE),
                                  0.5 * mean(Resisters$k48a, na.rm = TRUE),
                                  mean(Resisters$k57x, na.rm = TRUE)))

ev_df4 <- data.frame(Class = "Rejecters",
                     Variable = 1:6,
                     Expected = c(mean(Rejecters$k32ax, na.rm = TRUE),
                                  mean(Rejecters$k32bx, na.rm = TRUE),
                                  mean(Rejecters$k32cx, na.rm = TRUE),
                                  mean(Rejecters$k32dx, na.rm = TRUE),
                                  0.5 * mean(Rejecters$k48a, na.rm = TRUE),
                                  mean(Rejecters$k57x, na.rm = TRUE)))

# Combine all data frames
ev_plot <- rbind(ev_df1, ev_df2, ev_df3, ev_df4)

# Define predictor variable labels
pred_vars <- c("D1:Usage", "D1:Search", "D1:Privacy", "D1:Learning", "Inst. Trust*", "AI Threat")

# Convert variables to factor with proper labels
ev_plot$Variable <- factor(ev_plot$Variable, levels = 1:6, labels = pred_vars)
ev_plot$Class <- factor(ev_plot$Class, levels = c("Adopters", "Cautious", "Resisters", "Rejecters"))

# Plot
ggplot(data = ev_plot, aes(x = Variable, y = Expected, colour = Class, group = Class)) +
  geom_point(shape = 2, size = 2) +
  geom_line() +
  labs(x = "Predictor Variables", 
       y = "Expected Values",
       color = "Construct",
       title = "Expected Values by Latent Class",
       subtitle = "*Institutional trust scaled (0-5) for comparative readability") +
  scale_color_manual(values = paletteer::paletteer_d("nbapalettes::knicks_holiday", n = 4))+
  theme_pubr() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size=8))+
  plot_annotation(caption = "Data from FSD3994 ISSP 2024: Digital Societies: Finnish Data")
```


```{r Expected values alt, fig.cap="Expected values of other variables by latent class", fig.height=6}
# Create dataframes of expected values for each class
evx_df1 <- data.frame(Class = "Adopters",
                     Variable = 1:6,
                     Expected = c(mean(Adopters$k34ax, na.rm = TRUE),
                                  mean(Adopters$k34bx, na.rm = TRUE),
                                  0.5 * mean(Adopters$k46, na.rm = TRUE),
                                  0.5 * mean(Adopters$k47, na.rm = TRUE),
                                  0.5 * mean(Adopters$k48b, na.rm = TRUE),
                                  0.5 * mean(Adopters$k62, na.rm = TRUE)))

evx_df2 <- data.frame(Class = "Cautious",
                     Variable = 1:6,
                     Expected = c(mean(Cautious$k34ax, na.rm = TRUE),
                                  mean(Cautious$k34bx, na.rm = TRUE),
                                  0.5 * mean(Cautious$k46, na.rm = TRUE),
                                  0.5 * mean(Cautious$k47, na.rm = TRUE),
                                  0.5 * mean(Cautious$k48b, na.rm = TRUE),
                                  0.5 * mean(Cautious$k62, na.rm = TRUE)))

evx_df3 <- data.frame(Class = "Resisters",
                     Variable = 1:6,
                     Expected = c(mean(Resisters$k34ax, na.rm = TRUE),
                                  mean(Resisters$k34bx, na.rm = TRUE),
                                  0.5 * mean(Resisters$k46, na.rm = TRUE),
                                  0.5 * mean(Resisters$k47, na.rm = TRUE),
                                  0.5 * mean(Resisters$k48b, na.rm = TRUE),
                                  0.5 * mean(Resisters$k62, na.rm = TRUE)))

evx_df4 <- data.frame(Class = "Rejecters",
                     Variable = 1:6,
                     Expected = c(mean(Rejecters$k34ax, na.rm = TRUE),
                                  mean(Rejecters$k34bx, na.rm = TRUE),
                                  0.5 * mean(Rejecters$k46, na.rm = TRUE),
                                  0.5 * mean(Rejecters$k47, na.rm = TRUE),
                                  0.5 * mean(Rejecters$k48b, na.rm = TRUE),
                                  0.5 * mean(Rejecters$k62, na.rm = TRUE)))

# Combine all data frames
evx_plot <- rbind(evx_df1, evx_df2, evx_df3, evx_df4)

# Define predictor variable labels
predx_vars <- c("Prefers meeting online", "Lonely without internet", "Political leaning (Left=0)*", "Trust people in General*", "Trust Finnish courts*", "Self-assesed social level*")

# Convert variables to factor with proper labels
evx_plot$Variable <- factor(evx_plot$Variable, levels = 1:6, labels = predx_vars)
evx_plot$Class <- factor(evx_plot$Class, levels = c("Adopters", "Cautious", "Resisters", "Rejecters"))

# Plot
ggplot(data = evx_plot, aes(x = Variable, y = Expected, colour = Class, group = Class)) +
  geom_point(shape = 2, size = 2) +
  geom_line() +
  labs(x = "Predictor Variables", 
       y = "Expected Values",
       color = "Construct",
       title = "Expected Values by Latent Class",
       subtitle = "*Scaled (0-5) for comparative readability") +
  scale_color_manual(values = paletteer::paletteer_d("nbapalettes::knicks_holiday", n = 4))+
  theme_pubr() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8))+
  plot_annotation(caption = "Data from FSD3994 ISSP 2024: Digital Societies: Finnish Data")
```

# Indicator distributions by latent class

```{r anx class 1 predictor plots, fig.cap="Class 1 predictor distributions", fig.height=6}

ind_class1 <- wrap_plots(
  k32a_class1, k32b_class1, 
  k32c_class1, k32d_class1,
  k48a_class1, k57x_class1,
  ncol = 2
) + 
  plot_annotation(title = "Class 1 (Adopters)", 
                  caption = "Data from FSD3994 ISSP 2024: Digital Societies: Finnish Data")

ind_class1 + 
  theme(plot.margin = margin(10, 10, 10, 10))
```

```{r anx class 1 demo plots, fig.cap="Class 1 demographic distributions", fig.height=6}

demo_class1 <- wrap_plots(
  gender_class1, age_class1, 
  edu_class1,
  ncol = 1
) + 
  plot_annotation(title = "Class 1 (Adopters)", 
                  caption = "Data from FSD3994 ISSP 2024: Digital Societies: Finnish Data")

demo_class1 + 
  theme(plot.margin = margin(10, 10, 10, 10))

```

```{r anx class 1 ika stats, tab.cap="Summary statistics for age variable 'ika'"}
library(dplyr)

# Summary statistics for the variable "ika"
summary_ika_class1 <- Adopters %>%
  summarise(
    mean = mean(ika, na.rm = TRUE),
    sd = sd(ika, na.rm = TRUE),
    min = min(ika, na.rm = TRUE),
    Q1 = quantile(ika, 0.25, na.rm = TRUE),  
    Q3 = quantile(ika, 0.75, na.rm = TRUE),
    max = max(ika, na.rm = TRUE),
    na_count = sum(is.na(ika))
  )

ika_table_class1 <- summary_ika_class1 %>%
  gt() %>%
  tab_header(
    title = "Class 1 (Adopters) age summary statistics"
  ) %>%
  cols_label(
    mean = "Mean",
    sd = "Standard Deviation",
    min = "Minimum",
    Q1 = "Quartile 1",
    Q3 = "Quartile 3",
    max = "Maximum",
    na_count = "NA Count"
  ) %>%
  fmt_number(
    columns = c(mean, sd, min, max),
    decimals = 2
  ) %>%
  fmt_number(
    columns = c(na_count),
    decimals = 0
  )
ika_table_class1
```

```{r anx class 2 predictor plots, fig.cap="Class 2 predictor distributions", fig.height=6}

ind_class2 <- wrap_plots(
  k32a_class2, k32b_class2, 
  k32c_class2, k32d_class2,
  k48a_class2, k57x_class2,
  ncol = 2
) + 
  plot_annotation(title = "Class 2 (Cautious)", 
                  caption = "Data from FSD3994 ISSP 2024: Digital Societies: Finnish Data")

ind_class2 + 
  theme(plot.margin = margin(10, 10, 10, 10))
```


```{r anx class 2 demo plots, fig.cap="Class 2 demographic distributions", fig.height=6}

demo_class2 <- wrap_plots(
  gender_class2, age_class2, 
  edu_class2,
  ncol = 1
) + 
  plot_annotation(title = "Class 2 (Cautious)", 
                  caption = "Data from FSD3994 ISSP 2024: Digital Societies: Finnish Data")

demo_class2 + 
  theme(plot.margin = margin(10, 10, 10, 10))


```

```{r anx class 2 ika stats, tab.cap="Summary statistics for age variable 'ika'"}
library(dplyr)

# Summary statistics for the variable "ika"
summary_ika_class2 <- Cautious %>%
  summarise(
    mean = mean(ika, na.rm = TRUE),
    sd = sd(ika, na.rm = TRUE),
    min = min(ika, na.rm = TRUE),
    Q1 = quantile(ika, 0.25, na.rm = TRUE),  
    Q3 = quantile(ika, 0.75, na.rm = TRUE),
    max = max(ika, na.rm = TRUE),
    na_count = sum(is.na(ika))
  )

ika_table_class2 <- summary_ika_class2 %>%
  gt() %>%
  tab_header(
    title = "Class 2 (Cautious) age summary statistics"
  ) %>%
  cols_label(
    mean = "Mean",
    sd = "Standard Deviation",
    min = "Minimum",
    Q1 = "Quartile 1",
    Q3 = "Quartile 3",
    max = "Maximum",
    na_count = "NA Count"
  ) %>%
  fmt_number(
    columns = c(mean, sd, min, max),
    decimals = 2
  ) %>%
  fmt_number(
    columns = c(na_count),
    decimals = 0
  )
ika_table_class2
```


```{r anx class 3 predictor plots, fig.cap="Class 3 predictor distributions", fig.height=6}

ind_class3 <- wrap_plots(
  k32a_class3, k32b_class3, 
  k32c_class3, k32d_class3,
  k48a_class3, k57x_class3,
  ncol = 2
) + 
  plot_annotation(title = "Class 3 (Resisters)", 
                  caption = "Data from FSD3994 ISSP 2024: Digital Societies: Finnish Data")

ind_class3 + 
  theme(plot.margin = margin(10, 10, 10, 10))
```


```{r anx class 3 demo plots, fig.cap="Class 3 demographic distributions", fig.height=6}


demo_class3 <- wrap_plots(
  gender_class3, age_class3, 
  edu_class3,
  ncol = 1
) + 
  plot_annotation(title = "Class 3 (Resisters)", 
                  caption = "Data from FSD3994 ISSP 2024: Digital Societies: Finnish Data")

demo_class3 + 
  theme(plot.margin = margin(10, 10, 10, 10))


```

```{r anx class 3 ika stats, tab.cap="Summary statistics for age variable 'ika'"}
library(dplyr)

# Summary statistics for the variable "ika"
summary_ika_class3 <- Resisters %>%
  summarise(
    mean = mean(ika, na.rm = TRUE),
    sd = sd(ika, na.rm = TRUE),
    min = min(ika, na.rm = TRUE),
    Q1 = quantile(ika, 0.25, na.rm = TRUE),  
    Q3 = quantile(ika, 0.75, na.rm = TRUE),
    max = max(ika, na.rm = TRUE),
    na_count = sum(is.na(ika))
  )

ika_table_class3 <- summary_ika_class3 %>%
  gt() %>%
  tab_header(
    title = "Class 3 (Resisters) age summary statistics"
  ) %>%
  cols_label(
    mean = "Mean",
    sd = "Standard Deviation",
    min = "Minimum",
    Q1 = "Quartile 1",
    Q3 = "Quartile 3",
    max = "Maximum",
    na_count = "NA Count"
  ) %>%
  fmt_number(
    columns = c(mean, sd, min, max),
    decimals = 2
  ) %>%
  fmt_number(
    columns = c(na_count),
    decimals = 0
  )
ika_table_class3
```


```{r anx class 4 predictor plots, fig.cap="Class 4 predictor distributions", fig.height=6}

ind_class4 <- wrap_plots(
  k32a_class4, k32b_class4, 
  k32c_class4, k32d_class4,
  k48a_class4, k57x_class4,
  ncol = 2
) + 
  plot_annotation(title = "Class 4 (Rejecters)", 
                  caption = "Data from FSD3994 ISSP 2024: Digital Societies: Finnish Data")

ind_class4 + 
  theme(plot.margin = margin(10, 10, 10, 10))
```


```{r anx class 4 demo plots, fig.cap="Class 4 demographic distributions", fig.height=6}

demo_class4 <- wrap_plots(
  gender_class4, age_class4, 
  edu_class4,
  ncol = 1
) + 
  plot_annotation(title = "Class 4 (Rejecters)", 
                  caption = "Data from FSD3994 ISSP 2024: Digital Societies: Finnish Data")

demo_class4 + 
  theme(plot.margin = margin(10, 10, 10, 10))


```

```{r anx class 4 ika stats, tab.cap="Summary statistics for age variable 'ika'"}
library(dplyr)

# Summary statistics for the variable "ika"
summary_ika_class4 <- Rejecters %>%
  summarise(
    mean = mean(ika, na.rm = TRUE),
    sd = sd(ika, na.rm = TRUE),
    min = min(ika, na.rm = TRUE),
    Q1 = quantile(ika, 0.25, na.rm = TRUE),  
    Q3 = quantile(ika, 0.75, na.rm = TRUE),
    max = max(ika, na.rm = TRUE),
    na_count = sum(is.na(ika))
  )

ika_table_class4 <- summary_ika_class4 %>%
  gt() %>%
  tab_header(
    title = "Class 4 (Rejecters) age summary statistics"
  ) %>%
  cols_label(
    mean = "Mean",
    sd = "Standard Deviation",
    min = "Minimum",
    Q1 = "Quartile 1",
    Q3 = "Quartile 3",
    max = "Maximum",
    na_count = "NA Count"
  ) %>%
  fmt_number(
    columns = c(mean, sd, min, max),
    decimals = 2
  ) %>%
  fmt_number(
    columns = c(na_count),
    decimals = 0
  )
ika_table_class4
```

\clearpage

# GLM Regression: Model of Full Sample 

## A1 Models: Comfort with AI Communication

```{r AR1 GLM Regression model 1}
# GLM with a Gaussian family (AR1 is continuous)
AR1_glm1 <- glm(A1 ~ D1, data = AiRf, family = gaussian())

# Summary of the GLM
summary(AR1_glm1)
```

```{r AR1 GLM Regression model 2}
# GLM with a Gaussian family (AR1 is continuous)
AR1_glm2 <- glm(A1 ~ D1 + k48a, data = AiRf, family = gaussian())

# Summary of the GLM
summary(AR1_glm2)
```

```{r AR1 GLM Regression model 3}
# GLM with a Gaussian family (AR1 is continuous)
AR1_glm3 <- glm(A1 ~ D1 + k48a + k57x, data = AiRf, family = gaussian())

# Summary of the GLM
summary(AR1_glm3)
```

```{r AR1 GLM Regression model 4}
# GLM with a Gaussian family (AR1 is continuous)
AR1_glm4 <- glm(A1 ~ D1 + k48a + k57x + ika + k1 + k4, data = AiRf, family = gaussian())

# Summary of the GLM
summary(AR1_glm4)
```

```{r AR1 Linear model table}
# Create a summary table for the models
stargazer(AR1_glm1, AR1_glm2, AR1_glm3, AR1_glm4,
          type = "latex",
          title = "Summary of GLM Models",
          dep.var.labels = "Outcome Variable (A1)",
          column.labels = c("Model 1", "Model 2", "Model 3", "Model 4"),
          model.names = FALSE,
          single.row = TRUE,
          digits = 3,
          star.char = c("*", "**", "***"),
          star.cutoffs = c(0.05, 0.01, 0.001),
          covariate.labels = c("DL1", "k48a", "k57x", "ika", "k1", "k4"),
          omit.stat = c("f", "ser"))  
```


```{r A1 Test residuals model 4, fig.cap="Analysis of A1 GLM model (Full Sample)"}
# Set up the plotting area
par(mfrow = c(2, 2)) 

plot(AR1_glm4, main= "A1 Model 4")
```

## A2 Models: Comfort with AI Automation 

```{r AR2 GLM Regression model 1}
# GLM with a Gaussian family (AR2 is continuous)
AR2_glm1 <- glm(A2 ~ D1, data = AiRf, family = gaussian())

# Summary of the GLM
summary(AR2_glm1)
```

```{r AR2 GLM Regression model 2}
# GLM with a Gaussian family (AR2 is continuous)
AR2_glm2 <- glm(A2 ~ D1 + k48a, data = AiRf, family = gaussian())

# Summary of the GLM
summary(AR1_glm2)
```

```{r AR2 GLM Regression model 3}
# GLM with a Gaussian family (AR2 is continuous)
AR2_glm3 <- glm(A2 ~ D1 + k48a + k57x, data = AiRf, family = gaussian())

# Summary of the GLM
summary(AR1_glm3)
```

```{r AR2 GLM Regression model 4}
# GLM with a Gaussian family (AR2 is continuous)
AR2_glm4 <- glm(A2 ~ D1 + k48a + k57x + ika + k1 + k4, data = AiRf, family = gaussian())

# Summary of the GLM
summary(AR2_glm4)
```

```{r AR2 Linear model table}
# Create a summary table for the models
stargazer(AR2_glm1, AR2_glm2, AR2_glm3, AR2_glm4,
          type = "latex",
          title = "Summary of GLM Models",
          dep.var.labels = "Outcome Variable (AR1)",
          column.labels = c("Model 1", "Model 2", "Model 3"),
          model.names = FALSE,
          single.row = TRUE,
          digits = 3,
          star.char = c("*", "**", "***"),
          star.cutoffs = c(0.05, 0.01, 0.001),
          covariate.labels = c("DL1", "k48a", "k57x", "ika", "k1", "k4")
          )  
```


```{r A2 Test residuals model 4, fig.cap="Analysis of A2 GLM model (Full Sample)"}
# Set up the plotting area
par(mfrow = c(2, 2)) 

plot(AR2_glm4, main= "A2 Model 4")
```


# GLM Regression: Effect of Excluding Resister Class 

```{r Test sample after excluding Resisters}

TAMpop <- rbind(Adopters, Cautious, Rejecters)

# AI receptivity (comfort) subset
EFAAIx <- subset(TAMpop, select= c(k58a,k58b,k58c,k58d,k58e))

EFAAIx <- EFAAIx%>% rename(Medical = k58a,
                         Driving = k58b, 
                         Uncertainty = k58c,
                         Interaction = k58d,
                         Reliability = k58e)
```

```{r AI comfort corr resist excluded, fig.cap="Correlation of AI comfort indicators (Resister class excluded))"}
# Calculate the correlation matrix
AI_matx <- cor(EFAAIx, method = "spearman")


# plot the matrix
corrplot(AI_matx, method = "color", 
         type = "full",  
         order = "original", 
         tl.col = "black",
         tl.cex = .8,
         diag = TRUE,
         addgrid.col = "cadetblue",
         addrect = 1,
         number.font = 1,
         tl.srt = 45,
         addCoef.col = "black",  
         number.cex = 1,      
         title = "AI Correlation matrix (resisters excluded)", 
         mar=c(0,0,2,0))         


```


```{r AIR parallel analysis excl resist, fig.cap="Parallel analysis test for AI receptivity (Resisters excluded)"}

# Parallel analysis test for number of factors
fa.parallel(EFAAIx,
            fa = "fa",
            main="PA: Resiters Excluded",
            n.iter = 10,
            correct=.5, 
            sqrt=TRUE)

```

```{r EFA excl resist}

fa_aix <- fa(EFAAIx, 
              nfactors = 2, 
              rotate = "Promax",
              cor = "spearman",
              fm ="uls", 
              scores = "regression")
print(fa_aix)
```

```{r AI EFA diag excl resist, fig.cap="AI comfort factor structure (Resisters excluded)"}
# Factor loading diagram digital literacy
colnames(fa_aix$loadings) <- c("B1", "B2")
fa.diagram(fa_aix, 
           marg=c(1,1,1,1),
           main="",
           l.cex = 0.8,
           e.size = .1,
           cut = 0.3)

```

```{r AI Factor loadings variables excl resist}
# AI Factor loading variable
fact_score_aix <- as.data.frame(fa_aix$scores)
fact_score_aix <- fact_score_aix %>% rename(B1 = ULS1)
fact_score_aix <- fact_score_aix %>% rename(B2 = ULS2)
fact_score_aix$B1 <- set_label(fact_score_aix$B1, label = "B1 : Communication")
fact_score_aix$B2 <- set_label(fact_score_aix$B2, label = "B2 : Automation")

AiRx <- cbind(TAMpop, fact_score_aix)

```

```{r EFA B1 B2 loadings, fig.cap="AI Comfort factor analysis loadings (Resisters excluded)"}
# Extract loadings and communalities
loadings1_aix <- fa_aix$loadings[, 1]  
loadings2_aix <- fa_aix$loadings[, 2]  
communalities_aix <- fa_aix$communality  

# Data frame for the table
df_aix <- data.frame(
  Indicator = c("Medical", "Driving", "Uncertainty", "Interaction", "Reliability"),
  Variable = rownames(fa_ai$loadings),  
  B1 = loadings1_aix,
  B2 = loadings2_aix,
  Communalities = communalities_aix          
)

# gt table
ai_tablex <- df_aix %>%
  gt() %>%
  tab_header(
    title = "AI Comfort EFA Results (Resisters Excluded)",
    subtitle = "Factor Loadings and Communalities"
  ) %>%
  cols_label(
    Indicator = "Comfort Measure",
    Variable = "Variable",
    B1 = "B1",
    B2 = "B2",
    Communalities = "h² Communalities"
  ) %>%
  fmt_number(
    columns = c(B1, B2, Communalities),
    decimals = 3  
  )
ai_tablex

```

## B1 Models: Comfort with AI Communication 

```{r B1 GLM Regression model 1}
# GLM with a Gaussian family (AR1 is continuous)
B1_glm1 <- glm(B1 ~ D1, data = AiRx, family = gaussian())

# Summary of the GLM
summary(B1_glm1)
```

```{r B1 GLM Regression model 2}
# GLM with a Gaussian family (AR1 is continuous)
B1_glm2 <- glm(B1 ~ D1 + k48a, data = AiRx, family = gaussian())

# Summary of the GLM
summary(B1_glm2)
```

```{r B1 GLM Regression model 3}
# GLM with a Gaussian family (AR1 is continuous)
B1_glm3 <- glm(B1 ~ D1 + k48a + k57x, data = AiRx, family = gaussian())

# Summary of the GLM
summary(B1_glm3)
```

```{r B1 GLM Regression model 4}
# GLM with a Gaussian family (AR1 is continuous)
B1_glm4 <- glm(B1 ~ D1 + k48a + k57x + ika + k1 + k4, data = AiRx, family = gaussian())

# Summary of the GLM
summary(B1_glm4)
```

```{r B1 Linear model table}
# Create a summary table for the models
stargazer(B1_glm1, B1_glm2, B1_glm3, B1_glm4,
          type = "latex",
          title = "Summary of B1 GLM Models (Resisters Excluded)",
          dep.var.labels = "Outcome Variable (B1)",
          column.labels = c("Model 1", "Model 2", "Model 3", "Model 4"),
          model.names = FALSE,
          single.row = TRUE,
          digits = 3,
          star.char = c("*", "**", "***"),
          star.cutoffs = c(0.05, 0.01, 0.001),
          covariate.labels = c("D1", "k48a", "k57x", "ika", "k1", "k4"),
          omit.stat = c("f", "ser"))  
```


```{r B1 Test residuals model 4, fig.cap="Analysis of B1 GLM model (Resisters excluded)"}
# Set up the plotting area
par(mfrow = c(2, 2)) 

plot(B1_glm4, main= "Model 4 (B1)")
```

## B2 Models: Comfort with AI Automation 

```{r B2 GLM Regression model 1}
# GLM with a Gaussian family (AR1 is continuous)
B2_glm1 <- glm(B2 ~ D1, data = AiRx, family = gaussian())

# Summary of the GLM
summary(B2_glm1)
```

```{r B2 GLM Regression model 2}
# GLM with a Gaussian family (AR1 is continuous)
B2_glm2 <- glm(B2 ~ D1 + k48a, data = AiRx, family = gaussian())

# Summary of the GLM
summary(B2_glm2)
```

```{r B2 GLM Regression model 3}
# GLM with a Gaussian family (AR1 is continuous)
B2_glm3 <- glm(B2 ~ D1 + k48a + k57x, data = AiRx, family = gaussian())

# Summary of the GLM
summary(B2_glm3)
```

```{r B2 GLM Regression model 4}
# GLM with a Gaussian family (AR1 is continuous)
B2_glm4 <- glm(B2 ~ D1 + k48a + k57x + ika + k1 + k4, data = AiRx, family = gaussian())

# Summary of the GLM
summary(B2_glm4)
```

```{r B2 Linear model table}
# Create a summary table for the models
stargazer(B2_glm1, B2_glm2, B2_glm3, B2_glm4,
          type = "latex",
          title = "Summary of GLM Models (Resisters Excluded)",
          dep.var.labels = "Outcome Variable (B2)",
          column.labels = c("Model 1", "Model 2", "Model 3", "Model 4"),
          model.names = FALSE,
          single.row = TRUE,
          digits = 3,
          star.char = c("*", "**", "***"),
          star.cutoffs = c(0.05, 0.01, 0.001),
          covariate.labels = c("D1", "k48a", "k57x", "ika", "k1", "k4"),
          omit.stat = c("f", "ser"))  
```


```{r B2 Test residuals model 4, fig.cap="Analysis of B2 GLM model (Resisters excluded)"}
# Set up the plotting area
par(mfrow = c(2, 2)) 

plot(B2_glm4, main= "Model 4 (B2)")
```



```{r B1 predictor plots ecl resist}

A1D1plot <- ggplot(
  data = AiRx,
  mapping = aes(x =D1, y = B1)) +
  geom_point() +
  #geom_boxplot(fill="steelblue3", alpha=0.6, varwidth=TRUE, staplewidth = 0.2, coef=1.5) +
  labs(x = "D1: Digital Efficacy") +
  geom_smooth(method = "lm") +
  theme_pubr()

A1k48plot <- ggplot(
  data = AiRx,
  mapping = aes(x =k48a_f, y = B1)) +
  geom_point(color="steelblue4", alpha=0.5) +
  geom_boxplot(fill="steelblue3", alpha=0.8, varwidth=TRUE, staplewidth = 0.2, coef=1.5) +
  labs(x = "Institutional Trust") +
  theme_pubr()

A1k57plot <- ggplot(
  data = AiRx,
  mapping = aes(x =k57x_f, y = B1)) +
  geom_point(color="steelblue4", alpha=0.5) +
  geom_boxplot(fill="steelblue3", alpha=0.8, varwidth=TRUE, staplewidth = 0.2, coef=1.5) +
  labs(x = "Personal Threat") +
  theme_pubr()

A1ikaplot <- ggplot(
  data = AiRx,
  mapping = aes(x =ika, y = B1)) +
  geom_point(color="steelblue4", alpha=0.5) +
    geom_smooth(method = "lm") +
  labs(x = "Age") +
  theme_pubr()

A1k1plot <- ggplot(
  data = AiRx,
  mapping = aes(x =k1_f, y = B1)) +
  geom_point(color="steelblue4", alpha=0.5) +
  geom_boxplot(fill="steelblue3", alpha=0.8, varwidth=TRUE, staplewidth = 0.2, coef=1.5) +
  labs(x = "Gender") +
  theme_pubr()

A1k4plot <- ggplot(
  data = AiRx,
  mapping = aes(x =k4_f, y = B1)) +
  geom_point(color="steelblue4", alpha=0.5) +
  geom_boxplot(fill="steelblue3", alpha=0.8, varwidth=TRUE, staplewidth = 0.2, coef=1.5) +
  labs(x = "Education Level") +
  theme_pubr()

```

```{r B2 predictor plots excl resist}

A2D1plot <- ggplot(
  data = AiRx,
  mapping = aes(x =D1, y = B2)) +
  geom_point() +
  #geom_boxplot(fill="steelblue3", alpha=0.6, varwidth=TRUE, staplewidth = 0.2, coef=1.5) +
  labs(x = "D1: Digital Efficacy") +
  geom_smooth(method = "lm") +
  theme_pubr()

A2k48plot <- ggplot(
  data = AiRx,
  mapping = aes(x =k48a_f, y = B2)) +
  geom_point(color="steelblue4", alpha=0.5) +
  geom_boxplot(fill="steelblue3", alpha=0.8, varwidth=TRUE, staplewidth = 0.2, coef=1.5) +
  labs(x = "Institutional Trust") +
  theme_pubr()

A2k57plot <- ggplot(
  data = AiRx,
  mapping = aes(x =k57x_f, y = B2)) +
  geom_point(color="steelblue4", alpha=0.5) +
  geom_boxplot(fill="steelblue3", alpha=0.8, varwidth=TRUE, staplewidth = 0.2, coef=1.5) +
  labs(x = "Personal Threat") +
  theme_pubr()

A2ikaplot <- ggplot(
  data = AiRx,
  mapping = aes(x =ika, y = B2)) +
  geom_point(color="steelblue4", alpha=0.5) +
    geom_smooth(method = "lm") +
  labs(x = "Age") +
  theme_pubr()

A2k1plot <- ggplot(
  data = AiRx,
  mapping = aes(x =k1_f, y = B2)) +
  geom_point(color="steelblue4", alpha=0.5) +
  geom_boxplot(fill="steelblue3", alpha=0.8, varwidth=TRUE, staplewidth = 0.2, coef=1.5) +
  labs(x = "Gender") +
  theme_pubr()

A2k4plot <- ggplot(
  data = AiRx,
  mapping = aes(x =k4_f, y = B2)) +
  geom_point(color="steelblue4", alpha=0.5) +
  geom_boxplot(fill="steelblue3", alpha=0.8, varwidth=TRUE, staplewidth = 0.2, coef=1.5) +
  labs(x = "Education Level") +
  theme_pubr()

```

```{r B1 plots ecl resist, fig.cap="B1: AI Comfort (Communication) predictor distributions", fig.height=6, message=FALSE}

a1_preds <- wrap_plots(
  A1D1plot, A1k48plot, 
  A1k57plot, A1ikaplot,
  A1k1plot, A1k4plot,
  ncol = 2
) + 
  plot_annotation(title = "Comfort with AI (Communication Factor)", 
                  caption = "Data from FSD3994 ISSP 2024: Digital Societies: Finnish Data")

a1_preds + 
  theme(plot.margin = margin(10, 10, 10, 10))

```

```{r B2 plots ecl resist, fig.cap="B2: AI Comfort (Automation) predictor distributions", fig.height=6, message=FALSE}

a2_preds <- wrap_plots(
  A2D1plot, A2k48plot, 
  A2k57plot, A2ikaplot,
  A2k1plot, A2k4plot,
  ncol = 2
) + 
  plot_annotation(title = "Comfort with AI (Automation Factor)", 
                  caption = "Data from FSD3994 ISSP 2024: Digital Societies: Finnish Data")

a2_preds + 
  theme(plot.margin = margin(10, 10, 10, 10))

```

